---
title: "1_Practical; Data Exploration"
author: "Ryan Greeuup 17805315"
date: "1 August 2019"
output:
  html_document:
    code_folding: hide
    keep_md: yes
    theme: flatly
    toc: yes
    toc_depth: 4
    toc_float: yes
  pdf_document:
    keep_tex: yes
    toc: yes
always_allow_html: yes
---
\fontsize{12}{22}
\fontfamily{cmss}
\selectfont

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if(require('pacman')){
  library('pacman')
}else{
  install.packages('pacman')
  library('pacman')
}

pacman::p_load(ggplot2, rmarkdown, shiny, ISLR, class, BiocManager, corrplot, plotly)
  #Mass isn't available for R 3.5...
```


# 1_Exercises (Wk 1 - Introduction to Data Science) 
## Exercises (Part 1) {.tabset}
For `iris`, `heart` and `groceries` data sets:

1. Explore the variables
2. List the variables and classify as quantitative or qualitative
3. Provide a Research Question and identify a target variable
4. Identify whether or not this would exemplify supervised or unsupervised learning.

### `iris` Data Set
```{r, collapse=TRUE}
iris <- read.csv(file = "Datasets/iris.csv", header = TRUE, sep = ",")
```
#### (1) Explore the varibles
In order to explore the variables use `str()`, `summary()` and `head()`
```{r, collapse=FALSE}
head(iris,4)
str(iris)
summary(iris)

```

#### (2) List the Variables
In this case the variables are:

|Variable Name | Type of Data | Measurement |
|--------------|--------------|-------------|
| Type, (presumably species)         | Categorical  | The type of flower (i.e. species)|
| PW, (petal width) | Quantitative | Linear Distance|
| SW, (sepal width) | Quantitative | Linear Distance|
| PL, (petal length) | Quantitative | Linear Distance|
| SL, (sepal length) | Quantitative | Linear Distance|

#### (3) provide a research Question
A possible research question could be:

* Can plant species be the Sepal Length and Petal Width.

#### (4) Is this Supervised or Unsupervised Learning
This research question would be an example of **supervised learning** because the plant species are known and the model can be trained using already-known output values.

### `heart` dataset

```{r}
heart <- read.csv(file = "Datasets/heart.csv", header = TRUE, sep = ",")
```
#### (1) Explore the varibles
In order to explore the variables use `str()`, `summary()` and `head()`
```{r}
head(heart,4)
str(heart)
summary(heart)

```

#### (2) List the Variables
In this case the variables are:

|Variable Name | Type of Data | Measurement |
|--------------|--------------|-------------|
| X  | Categorical ($\mathbb{N}$) | presumably observation number |
| Age  | Quantitative | The age of the individual|
| Sex  | Categorical| The individuals gender |
| Chestpain  |  Categorical | A classification of the type of chest pain |
| RestBP  | Quantitative | A measurement of Sys. Blood Pressure at rest |
| Chol   | Quantitative| Cholestrol levels |
| Fbs   | Categorical | An indicator of whether or not Fasting Blood Sugar is above a threshold|
| RestECG  | Categorical | An indicator of the type ECG result  |
| MaxHR  | Quantitative | A measurement of the maximum Heart Rate|
| ExAng  | Categorical | An indicator of whether or not this individual suffered exercise induced angina|
| oldpeak  | quantitative |A measurement of ECG change induced by exercise |
| slope  | categorical | An indicator of the slope of the ST segment of an ECG graph |
| Ca  | categorical (because it exists in $(\mathbb{N})$| An indicator of how many of the three major blood vessels are revealed by fluroscopy |
| AHD  |  categorical | An indicator of whether or not the individual suffered  Atherosclerotic Heart Disease |

#### (3) provide a research Question
A possible research question could be:

* does MaxHR predict  Atherosclerotic Heart Disease independent of age?

#### (4) Is this Supervised or Unsupervised Learning
This research question would be an example of **supervised learning** because the incidence of AHD are known and the model can be trained using already-known output values.


### `groceries` dataset

```{r}
groceries <- read.csv(file = "Datasets/groceries.csv", header = TRUE, sep = ",")
```
#### (1) Explore the varibles
In order to explore the variables use `str()`, `summary()` and `head()`
```{r}
head(groceries[,1:5], 4)
  #Restrict the columns of groceries to fit on the page
str(groceries[,1:5])
summary(groceries[,1:5])

```

#### (2) List the Variables
It is necessary to see whether or not the input value is 1/0 or a number, we can check this by doing:

```{r}
if(sum(groceries>1)==0){
  print("The values are Boolean")
} else{
  print("The values could be categorical or continuous")
}
```

In this case the variables are:

|Variable Name | Type of Data | Measurement |
|--------------|--------------|-------------|
| food item  | categorical | whether or not the item needs to be purchased|

The subsequent observations (rows) could represent the need for groceries at each week.

#### (3) provide a research Question
A possible research question could be:

* Are certain food items more common at during holiday periods,
    + so for example is consumption of processed meats more common, this could be a public health enquiry.

##### (4) Is this Supervised or Unsupervised Learning
This research question would be an example of **unsupervised learning** because the pattern of food consumption is not known and the algorithm must 'learn' what the patterns are.


