---
title: "Exercise 11.2"
output: 
  html_document: 
    toc: yes
  html_notebook: 
    highlight: tango
    theme: journal
  pdf_document: 
    toc: yes
  word_document: 
    toc: yes
---

# Preamble
## Import Library

The following code will automatically install packages that are 
not already installed and load them into the library:

```{r}
if(require('pacman')){
  library('pacman')
}else{
  install.packages('pacman')
  library('pacman')
}

pacman::p_load(sp, EnvStats, gstat, ggplot2, rmarkdown)
```

## Create Assignments

Note that  Benthic.df is included in the *EnvStats* package and is not a 
data frame but an *sp* class, which is very similar to a data frame

```{r}
#Assignments
benthic <- Benthic.df
lat <- benthic$Latitude
lon <- benthic$Longitude
```

#Inspect the Data Set

```{r}
head(benthic, 3)
```

#Asign Co-ordinates
The benthic data set is not a mere data frame, it is an `sp` object,
which is very similar to a data frame.
It is necessary to set the co-ordinates of this object though:
```{r}
coordinates(benthic) = ~Longitude + Latitude
```

#Create a Directional variogram
A variogram is a description of the structure of spacial data,
first this needs to be created before anything can 
be plotted or modelled:


Because a directional variogram is still of the class `variogramModel`, 
it is not possible (to the best of my knowledge) to use `ggplot2`
in order to plot it.

The `alpha` option allows a vector of directional degrees to be specified, in this case, 22.5, 67.5, 112.5 and 157.5.

```{r}
benthic.dvg <- variogram(Index ~ 1,
                         data=benthic, alpha=c(22.5, 67.5, 112.5, 157.5))
benthic.dvg.fit <- vgm(1,"Sph", 0.5,1,anis=c(30,0.4))
plot(benthic.dvg, benthic.dvg.fit,
     main="Figure 11.5  Directional variograms for Benthic Index")
```



















