---
title: "Assessment 2"
author: "Ryan Greenup 17805315"
date: "19 May 2020"
output:
  html_document:
    code_folding: hide
    keep_md: yes
    theme: flatly
    toc: yes
    toc_depth: 4
    toc_float: no
  pdf_document: 
    toc: yes
always_allow_html: yes
  ##Shiny can be good but {.tabset} will be more compatible with PDF
    ##but you can submit HTML in turnitin so it doesn't really matter.
    
    ##If a floating toc is used in the document only use {.tabset} on more or less copy/pasted 
        #sections with different datasets
---

<!--  https://github.com/rstudio/rmarkdown/issues/1453#issuecomment-425327570 How to use echo       --> 

```{r setup, include=FALSE, include = FALSE, results = "hide", eval = TRUE}
knitr::opts_chunk$set(echo = TRUE)

## Preamble
# setwd("~/Dropbox/Notes/DataSci/ThinkingAboutData/Assessment/")
## Install Pacman
load.pac <- function() {

  if(require("pacman")){
    library(pacman)
  }else{
    install.packages("pacman")
    library(pacman)
  }
  
  ## Install packages
    pacman::p_load(xts, sp, gstat, ggplot2, rmarkdown, reshape2, ggmap,
                 parallel, dplyr, plotly, tidyverse, reticulate, UsingR, Rmpfr,
                 swirl, corrplot, gridExtra, mise, latex2exp, tree, rpart, lattice,
                 coin, primes, epitools, maps, clipr, ggmap, RColorBrewer, latex2exp)
  
  ## Clean up
   mise()
   
   ## Set Defaults
   select <- dplyr::select
   filter <- dplyr::filter
}

load.pac()
```


# Chloropleth map
## World Map
```{r}

covid <- read.csv("/home/ryan/Notes/DataSci/Visual_Analytics/Assessment2/owid-covid-data.csv")

covid$location[covid$location=="United States"] <- "USA"

# Only Extreme Examples -----------------------------------------------------
fatalprop <- aggregate(total_deaths_per_million ~ location, covid, max)
fatalprop <- fatalprop[order(-fatalprop$total_deaths_per_million),]
# fatalprop <- fatalprop[1:20,]

## Rename to facilitate joining with map
names(fatalprop) <- c("region", "total_deaths_per_million")

## San Marino will be shown by italy and this skews the results
## Belgium and San Marino are very hard to visualise from above
## They skew the rsults and so will be removed.
fatalprop <- fatalprop %>% filter(region!="San Marino")
fatalprop <- fatalprop %>% filter(region!="Belgium")

# Retrievethe map data
some.eu.maps <- map_data("world", region = fatalprop$location)
covid$total_deaths

# Compute the centroid as the mean longitude and lattitude
# Used as label coordinate for country's names
region_lab_df <- some.eu.maps %>%
  group_by(region) %>%
  summarise(long = mean(long), lat = mean(lat)) %>% 
    full_join(aggregate(total_deaths_per_million ~ region, fatalmap, mean))
# Manually Adjust US to be population Centre
region_lab_df[region_lab_df$region == "USA",]$long <- -92.47
region_lab_df[region_lab_df$region == "USA",]$lat <- 37.37

fatalmap <- left_join(fatalprop, some.eu.maps, by = "region")

wmp <-  ggplot(fatalmap, aes(map_id = region)) +
  geom_map(map = fatalmap,  color = "grey", aes(fill = total_deaths_per_million), lwd = 0.1, alpha = 0.3)+
  expand_limits(x = fatalmap$long, y = fatalmap$lat)+
  scale_fill_gradient(high = "darkred", low = "white") +
  guides(fill = guide_legend("Total Deaths \n per Million")) +
   # Change the colors of plot panel background to lightblue
   # and the color of grid lines to white
   theme(
     panel.background = element_rect(fill = "lightblue",
                                     colour = "lightblue",
                                     size = 0.5, linetype = "solid"),
     legend.position = c(0.6, 0.1),
     legend.direction = "horizontal",
     legend.background = element_rect(fill = "white", size = 0.1, colour = "darkblue", linetype = "solid")) +
   labs(x = "Longitude", y = "Latitude", title = TeX("Total Deaths Attributed to \\textit{COVID-19}")) 
  # geom_text(data = region_lab_df, aes(y = lat, x = long, label = region), size = 1)
wmp

```


### Add Bubble Chart
```{r}
wmp +
  scale_size_continuous(range = c(1, 9), name = "Total Number \n of Deaths") +
   guides(size = FALSE) +
   geom_point(data = region_lab_df, aes(y = lat, x = long, size = total_deaths_per_million), alpha = 0.5, col = "purple")
```

## Europe
### Set up

```{r}

# Only Extreme Examples -----------------------------------------------------
fatalprop <- aggregate(total_deaths_per_million ~ location, covid, max)
fatalprop <- fatalprop[order(-fatalprop$total_deaths_per_million),]
 

## Rename to facilitate joining with map
names(fatalprop) <- c("region", "total_deaths_per_million")

## San Marino will be shown by italy
 fatalprop <- fatalprop %>% filter(region!="San Marino")

names(fatalprop)

# Retrievethe map data
some.eu.maps <- map_data("world", region = fatalprop$location)
covid$total_deaths

# Compute the centroid as the mean longitude and lattitude
# Used as label coordinate for country's names
fatalmap <- left_join(fatalprop, some.eu.maps, by = "region")

fatalprop[order(fatalprop$total_deaths),]
## Filter out only Europe 
fatalmap <-  fatalmap %>% 
  filter(30 <  lat & lat < 65) %>% 
  filter(-30 <  long & long < 35) 

(region_lab_df <- fatalmap %>%
  dplyr::group_by(region) %>%
  dplyr::summarise(long = mean(long), lat = mean(lat)) %>% 
    full_join(aggregate(total_deaths_per_million ~ region, fatalmap, mean)))

aggregate(total_deaths_per_million ~ region, formula = mean, data = fatalmap)
names(fatalmap)
aggregate(formula = total_deaths_per_million ~ region, FUN = mean, data = fatalmap)

```


### ggplot2


```{r}
library(ggrepel)
ggplot(fatalmap, aes(map_id = region, label = region)) +
  geom_map(map = fatalmap,
           aes(fill = total_deaths_per_million),
           color = "white") +
  geom_point(data = region_lab_df, aes(y = lat, x = long, size = total_deaths_per_million), alpha = 0.45, colour = "blue", stroke = 1, fill = "white", shape = 21) +
  scale_size_continuous(range = c(1, 25), name = "Total Number \n of Deaths") +
  guides(size = FALSE) +
  expand_limits(x = fatalmap$long, y = fatalmap$lat) +
  scale_fill_viridis_c(option = "C") +
  scale_fill_gradient(high = "darkred", low = "white") +
  guides(fill = guide_legend("Total Deaths \n per Million")) +
  # Change the colors of plot panel background to lightblue
  # and the color of grid lines to white
  theme(
    panel.background = element_rect(
      fill = "lightblue",
      colour = "lightblue",
      size = 0.5,
      linetype = "solid"
    ),
    legend.position = c(0.1, 0.6),
    legend.direction = "vertical",
    legend.background = element_rect(
      fill = "white",
      size =
        1.1,
      colour = "darkblue",
      linetype = "solid"
    )
  ) +
  labs(
    x = "Longitude",
    y = "Latitude",
    title = TeX("Total Deaths Attributed to \\textit{COVID-19}")
  ) +
  geom_text_repel(
    data = region_lab_df,
    aes(y = lat, x = long, label = region),
    size = 2,
    col = "black",
    nudge_y = 0.7,
    nudge_x = -0.5,
    min.segment.length = 0.6,
    force = 2
  )

```


# Time Series
## Basic Plot
```{r}
cv <- as_tibble(covid)

cv <- cv %>% 
  mutate(date = as.Date(date)) 

interested_locations <- c("Australia", "USA", "Italy", "Germany", "Belgium", "United Kingdom", "New Zealand", "Japan", "China")
interested_locations <- c("Australia", "USA", "Italy", "Germany", "Belgium", "United Kingdom")

cv <- cv %>% 
  filter(location %in% interested_locations) %>% 
  filter(total_cases > 100)

ggplot(cv, aes(y = total_cases, x = date, col = location, group = location)) +
  geom_line() +
  labs(x = element_blank(), y = "Total Cases", title = "Total Cases over Time") + 
  theme_bw() +
  guides(col = guide_legend("Location"))
  
```

## Basic Plot
### Total Cases
```{r}
cv <- as_tibble(covid)
cv <- cv %>% 
  mutate(date = as.Date(date)) 

interested_locations <- c("Australia", "USA", "Italy", "Germany", "Belgium", "United Kingdom", "New Zealand", "Japan", "China")
interested_locations <- c("Australia", "USA", "Italy", "Germany", "China", "United Kingdom")

cv <- cv %>% 
  filter(location %in% interested_locations) %>% 
  filter(total_cases > 100) %>% 
  mutate(total_cases = log10(total_cases))


ggplot(fortify(cv) , aes(y = total_cases, x = date, col = location, group = location)) +
  geom_line()  + 
  theme_bw() +
  theme(axis.title.x = element_blank()) + 
  labs(y = "Total Number of Cases (Log-10 Scale)", title = "Log Scaled Cases over Time") +
  guides(col = guide_legend("Location"))
#  geom_smooth()
```

* Changing the cases relative to the population size shows that a much much smaller fraction of the Chinese Population contracted the disease compared to other countries with a relatively smaller population.

### Per Capita
```{r}
cv <- as_tibble(covid)
cv <- cv %>% 
  mutate(date = as.Date(date)) 

interested_locations <- c("Australia", "USA", "Italy", "Germany", "Belgium", "United Kingdom", "New Zealand", "Japan", "China")
interested_locations <- c("Australia", "USA", "Italy", "Germany", "China", "United Kingdom")

cv <- cv %>% 
  filter(location %in% interested_locations) %>% 
  filter(total_cases_per_million > 100) %>% 
  mutate(total_cases_per_million = log10(total_cases_per_million))


ggplot(fortify(cv) , aes(y = total_cases_per_million, x = date, col = location, group = location)) +
  geom_line()  + 
  theme_bw() +
  theme(axis.title.x = element_blank()) + 
  labs(y = "Total Number of Cases (Log-10 Scale)", title = "Log Scaled Countries over Time per Million") +
  guides(col = guide_legend("Location"))
#  geom_smooth()
```

### Days since first Infection

This is difficult to read so instead we will use the number of days since the 100th infection

```{r}
cv <- as_tibble(covid)
cv <- cv %>% 
  mutate(date = as.Date(date)) 
cv <- cv[order(cv$date),]

# interested_locations <- c("Australia", "USA", "Italy", "Germany", "Belgium", "United Kingdom", "New Zealand", "Japan", "China")
interested_locations <- c("Australia", "USA", "Italy", "Germany", "Russia", "South Korea", "United Kingdom")

cv <- cv %>% 
  filter(location %in% interested_locations) %>% 
  filter(total_cases_per_million > 1) %>% 
  mutate(total_cases_per_million = log10(total_cases_per_million)) %>% 
  dplyr::select(date, total_cases_per_million, location) %>% 
  pivot_wider(names_from = location, values_from = total_cases_per_million)

for (i in 2:ncol(cv)) {
  ## Slide the Columns up and put the NA at the end
cv[,i] <-   pull(cv, i) %>% 
  lead(cv[,i] %>%
         is.na() %>%
         sum())
 ## Replace the date with the number of days
cv$date <- seq_len(nrow(cv))
}

cv <- cv %>% 
 pivot_longer(names(cv)[-1], names_to = "location", values_to = "total_cases_per_million")


ggplot(cv , aes(y = total_cases_per_million, x = date, col = location, group = location)) +
  geom_point(alpha = 0.3)  + 
  geom_smooth() +
  theme_bw() +
  labs(y = "Total Number of Cases (Log-10 Scale)", title = "Log Scaled Total COVID-19 Cases per Million", x = TeX("Days since Case \\textit{#100}")) +
  guides(col = guide_legend("Location"))
#  geom_smooth()
```



```{r}
cv <- as_tibble(covid)
cv <- cv %>% 
  mutate(date = as.Date(date)) 
cv <- cv[order(cv$date),]

interested_locations <- c("Australia", "USA", "Italy", "Germany", "Belgium", "United Kingdom", "New Zealand", "Japan", "China")
# interested_locations <- c("Australia", "USA", "Italy", "Germany", "Russia", "South Korea", "United Kingdom")

cv <- cv %>% 
  filter(location %in% interested_locations) %>% 
  filter(total_cases_per_million > 1) %>% 
  mutate(total_deaths_per_million = log10(total_deaths_per_million)) %>% 
  dplyr::select(date, total_deaths_per_million, location) %>% 
  pivot_wider(names_from = location, values_from = total_deaths_per_million)

for (i in 2:ncol(cv)) {
  ## Slide the Columns up and put the NA at the end
cv[,i] <-   pull(cv, i) %>% 
  lead(cv[,i] %>%
         is.na() %>%
         sum())
 ## Replace the date with the number of days
cv$date <- seq_len(nrow(cv))
}

cv <- cv %>% 
 pivot_longer(names(cv)[-1], names_to = "location", values_to = "total_deaths_per_million")


ggplot(cv , aes(y = total_deaths_per_million, x = date)) +
  geom_point(alpha = 0.3, aes(col = location, group = location))  + 
  geom_smooth(aes(col = location, group = location)) +
  theme_bw() +
  labs(y = TeX("Total Number of deaths (log_{10} Scale)"), title = TeX("Log Scaled Total \\textit{COVID-19} deaths per Million"), x = TeX("Days since Case \\textit{#100}")) +
  guides(col = guide_legend("Location"))
#  geom_smooth()
covid
```


### Facet Grid

Now I want to combine all of these into one big facet_grid:

#### Only The Log Stuff


```{r}
interested_locations <- c("Australia", "USA", "Italy", "Germany", "Russia", "South Korea", "United Kingdom")

###### Number of Cases
cv <- as_tibble(covid)
cv <- cv %>% 
  mutate(date = as.Date(date)) 
cv <- cv[order(cv$date),]

cv <- cv %>% 
  filter(location %in% interested_locations) %>% 
  filter(total_cases_per_million > 1) %>% 
  mutate(total_cases_per_million = log10(total_cases_per_million)) %>% 
  dplyr::select(date, total_cases_per_million, location) %>% 
  pivot_wider(names_from = location, values_from = total_cases_per_million)

for (i in 2:ncol(cv)) {
  ## Slide the Columns up and put the NA at the end
cv[,i] <-   pull(cv, i) %>% 
  lead(cv[,i] %>%
         is.na() %>%
         sum())
 ## Replace the date with the number of days
cv$date <- seq_len(nrow(cv))
}

cv_cases_log <- cv %>% 
 pivot_longer(names(cv)[-1], names_to = "location", values_to = "value") %>% 
  add_column(subject = "cases") %>% 
  add_column(scale = "log")


### Number of deaths

cv <- as_tibble(covid)
cv <- cv %>% 
  mutate(date = as.Date(date)) 
cv <- cv[order(cv$date),]

cv <- cv %>% 
  filter(location %in% interested_locations) %>% 
  filter(total_cases_per_million > 1) %>% 
  mutate(total_deaths_per_million = log10(total_deaths_per_million)) %>% 
  dplyr::select(date, total_deaths_per_million, location) %>% 
  pivot_wider(names_from = location, values_from = total_deaths_per_million)

for (i in 2:ncol(cv)) {
  ## Slide the Columns up and put the NA at the end
cv[,i] <-   pull(cv, i) %>% 
  lead(cv[,i] %>%
         is.na() %>%
         sum())
 ## Replace the date with the number of days
cv$date <- seq_len(nrow(cv))
}

cv_deaths_log <- cv %>% 
 pivot_longer(names(cv)[-1], names_to = "location", values_to = "value") %>% 
  add_column(subject = "deaths") %>% 
  add_column(scale = "log")


### Number of Tests

cv <- as_tibble(covid)
cv <- cv %>% 
  mutate(date = as.Date(date)) 
cv <- cv[order(cv$date),]

cv <- cv %>% 
  filter(location %in% interested_locations) %>% 
  filter(total_cases_per_million > 1) %>% 
  mutate(total_tests_per_thousand = log10(total_tests_per_thousand)) %>% 
  dplyr::select(date, total_tests_per_thousand, location) %>% 
  mutate(total_tests_per_thousand = total_tests_per_thousand/1000) %>% 
  pivot_wider(names_from = location, values_from = total_tests_per_thousand)



for (i in 2:ncol(cv)) {
  ## Slide the Columns up and put the NA at the end
cv[,i] <-   pull(cv, i) %>% 
  lead(cv[,i] %>%
         is.na() %>%
         sum())
 ## Replace the date with the number of days
cv$date <- seq_len(nrow(cv))
}

cv_tests_log <- cv %>% 
 pivot_longer(names(cv)[-1], names_to = "location", values_to = "value") %>% 
  add_column(subject = "tests") %>% 
  add_column(scale = "log")


cv <- rbind(cv_cases_log, cv_deaths_log, cv_tests_log)

cv %>% 
  filter(subject == "deaths")

ggplot(cv , aes(y = value, x = date)) +
  geom_point(alpha = 0.4, aes(col = location))  + 
   geom_smooth(aes(col = location)) +
  theme_bw() +
  labs(y = TeX("Total Number of deaths (log_{10} Scale)"), title = TeX("Log Scaled Total \\textit{COVID-19} deaths per Million"), x = TeX("Days since Case \\textit{#100}")) +
  guides(col = guide_legend("Location")) +
  facet_grid(. ~ subject)

```

#### All Togehter
##### Define the Log Stuff

```{r}
interested_locations <- c("Australia", "USA", "Italy", "Germany", "Russia", "South Korea", "United Kingdom")

###### Number of Cases
cv <- as_tibble(covid)
cv <- cv %>% 
  mutate(date = as.Date(date)) 
cv <- cv[order(cv$date),]

cv <- cv %>% 
  filter(location %in% interested_locations) %>% 
  filter(total_cases > 1) %>% 
  mutate(total_cases_per_million = log10(total_cases_per_million)) %>% 
  dplyr::select(date, total_cases_per_million, location) %>% 
  pivot_wider(names_from = location, values_from = total_cases_per_million)

for (i in 2:ncol(cv)) {
  ## Slide the Columns up and put the NA at the end
cv[,i] <-   pull(cv, i) %>% 
  lead(cv[,i] %>%
         is.na() %>%
         sum())
 ## Replace the date with the number of days
cv$date <- seq_len(nrow(cv))
}

cv_cases_log <- cv %>% 
 pivot_longer(names(cv)[-1], names_to = "location", values_to = "value") %>% 
  add_column(subject = "No. of Cases") %>% 
  add_column(scale = "Log-10 Scale")


### Number of deaths

cv <- as_tibble(covid)
cv <- cv %>% 
  mutate(date = as.Date(date)) 
cv <- cv[order(cv$date),]

cv <- cv %>% 
  filter(location %in% interested_locations) %>% 
  filter(total_cases > 1) %>% 
   mutate(total_deaths_per_million = log10(total_deaths_per_million)) %>% 
  dplyr::select(date, total_deaths_per_million, location) %>% 
  pivot_wider(names_from = location, values_from = total_deaths_per_million)

for (i in 2:ncol(cv)) {
  ## Slide the Columns up and put the NA at the end
cv[,i] <-   pull(cv, i) %>% 
  lead(cv[,i] %>%
         is.na() %>%
         sum())
 ## Replace the date with the number of days
cv$date <- seq_len(nrow(cv))
}

cv_deaths_log <- cv %>% 
 pivot_longer(names(cv)[-1], names_to = "location", values_to = "value") %>% 
  add_column(subject = "No. of Deaths") %>% 
  add_column(scale = "Log-10 Scale")


### Number of Tests

cv <- as_tibble(covid)
cv <- cv %>% 
  mutate(date = as.Date(date)) 
cv <- cv[order(cv$date),]

cv <- cv %>% 
  filter(location %in% interested_locations) %>% 
  filter(total_cases > 1) %>% 
  mutate(total_tests_per_thousand = log10(total_tests_per_thousand)) %>% 
  dplyr::select(date, total_tests_per_thousand, location) %>% 
  pivot_wider(names_from = location, values_from = total_tests_per_thousand)



for (i in 2:ncol(cv)) {
  ## Slide the Columns up and put the NA at the end
cv[,i] <-   pull(cv, i) %>% 
  lead(cv[,i] %>%
         is.na() %>%
         sum())
 ## Replace the date with the number of days
cv$date <- seq_len(nrow(cv))
}

cv_tests_log <- cv %>% 
 pivot_longer(names(cv)[-1], names_to = "location", values_to = "value") %>% 
  add_column(subject = "No. of Tests") %>% 
  add_column(scale = "Log-10")


cv <- rbind(cv_cases_log, cv_deaths_log, cv_tests_log)

cv %>% 
  filter(subject == "deaths")

p_per_cap <- ggplot(cv , aes(y = value, x = date)) +
  geom_point(alpha = 0.3, aes(col = location))  + 
   geom_smooth(aes(col = location), size = 0.5) +
  theme_bw() +
  labs(y = TeX("Count (log_{10} Scale)"), title = TeX("log_{10} Scale; Value of \\textit{COVID-19} Statistics over Time"), x = TeX("Days since Case \\textit{#1}"), subtitle = "Counts Per Million of population") +
  guides(col = guide_legend("Location")) +
  facet_grid(rows = vars(subject), scales = "free_y")
p_per_cap
```


##### Define the raw Stuff

```{r}
interested_locations <- c("Australia", "USA", "Italy", "Germany", "Russia", "South Korea", "United Kingdom")

###### Number of Cases
cv <- as_tibble(covid)
cv <- cv %>% 
  mutate(date = as.Date(date)) 
cv <- cv[order(cv$date),]

cv <- cv %>% 
  filter(location %in% interested_locations) %>% 
  filter(total_cases > 1) %>% 
# mutate(total_cases = log10(total_cases)) %>% 
  dplyr::select(date, total_cases_per_million, location) %>% 
  pivot_wider(names_from = location, values_from = total_cases_per_million)

for (i in 2:ncol(cv)) {
  ## Slide the Columns up and put the NA at the end
cv[,i] <-   pull(cv, i) %>% 
  lead(cv[,i] %>%
         is.na() %>%
         sum())
 ## Replace the date with the number of days
cv$date <- seq_len(nrow(cv))
}

cv_cases_raw <- cv %>% 
 pivot_longer(names(cv)[-1], names_to = "location", values_to = "value") %>% 
  add_column(subject = "No. of Cases") %>% 
  add_column(scale = "Count")

### Number of deaths

cv <- as_tibble(covid)
cv <- cv %>% 
  mutate(date = as.Date(date)) 
cv <- cv[order(cv$date),]

cv <- cv %>% 
  filter(location %in% interested_locations) %>% 
  filter(total_cases > 1) %>% 
#  mutate(total_deaths = log10(total_deaths_)) %>% 
  dplyr::select(date, total_deaths_per_million, location) %>% 
  pivot_wider(names_from = location, values_from = total_deaths_per_million)

for (i in 2:ncol(cv)) {
  ## Slide the Columns up and put the NA at the end
cv[,i] <-   pull(cv, i) %>% 
  lead(cv[,i] %>%
         is.na() %>%
         sum())
 ## Replace the date with the number of days
cv$date <- seq_len(nrow(cv))
}

cv_deaths_raw <- cv %>% 
 pivot_longer(names(cv)[-1], names_to = "location", values_to = "value") %>% 
  add_column(subject = "No. of Deaths") %>% 
  add_column(scale = "Count")


### Number of Tests

cv <- as_tibble(covid)
cv <- cv %>% 
  mutate(date = as.Date(date)) 
cv <- cv[order(cv$date),]

cv <- cv %>% 
  filter(location %in% interested_locations) %>% 
  filter(total_cases > 1) %>% 
 # mutate(total_testsd = log10(total_testsd)) %>% 
  mutate(total_tests_per_thousand = total_tests_per_thousand/1000) %>% 
  dplyr::select(date, total_tests_per_thousand, location) %>% 
  pivot_wider(names_from = location, values_from = total_tests_per_thousand)



for (i in 2:ncol(cv)) {
  ## Slide the Columns up and put the NA at the end
cv[,i] <-   pull(cv, i) %>% 
  lead(cv[,i] %>%
         is.na() %>%
         sum())
 ## Replace the date with the number of days
cv$date <- seq_len(nrow(cv))
}

cv_tests_raw <- cv %>% 
 pivot_longer(names(cv)[-1], names_to = "location", values_to = "value") %>% 
  add_column(subject = "No. of Tests") %>% 
  add_column(scale = "Count")

cv <- rbind(cv_cases_raw, cv_deaths_raw, cv_tests_raw)

cv %>% 
  filter(subject == "deaths")

p_total <- ggplot(cv , aes(y = value, x = date)) +
  geom_point(alpha = 0.3, aes(col = location))  + 
   geom_smooth(aes(col = location), size = 0.5) +
  theme_bw() +
  labs(y = TeX("Total Count"), title = TeX("Total Count of \\textit{COVID-19} Statistics over Time"), x = TeX("Days since Case \\textit{#1}")) +
  guides(col = guide_legend("Location"), subtitle = "Per Million of Population") +
  facet_grid(rows = vars(subject), scales = "free_y")
p_total
```

##### Make the Combined Plot
```{r}
plots <- list(p_per_cap + guides(col = FALSE), p_total+ theme(legend.position="bottom") )
# plots <- list(p_per_cap + theme(legend.position="bottom"), p_total+ theme(legend.position="bottom") )
library(gridExtra)

gridExtra::grid.arrange(grobs = plots, layout_matrix = matrix(1:2, nrow = 1))

```

## Boxplots
```{r}
cv <- covid %>% 
  dplyr::filter(location %in% interested_locations)

ggplot(fortify(cv), aes(x = as.Date(date), y = new_cases_per_million, fill = location)) +
  geom_col(col = "grey") + 
  labs(x = "Date", y = "New Cases Per Million")
   theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme_bw()
  

```

