---
title: "The Marvel Universe as a Social Network"
csl: /home/ryan/Zotero/styles/nature.csl
output:
  html_document:
    df_print: paged
  html_notebook: null
bibliography: /home/ryan/Dropbox/Studies/Papers/references.bib
---

```{r, echo = FALSE, include=FALSE}
library(tidyverse)
library(plotly)
library(igraph)
library(reshape2)
library(mise)
mise()
```


# Introduction

In 2002 R. Alberich, J. Miro-Julia and F. Rossello investigated the relationship between characters in the Marvel Universe and compared this relational data to actual social networks [@alberich2002], the data set has been meed made available through *Kaggle* which will allow for further investigation [@zotero-400].

## Inspect the data

The data is provided in two formats, a list of characters and comic books in which they appear as well as a list of edges like so:

```{r}
h_net <- read.csv(file = "./hero-network.csv")
h_net[sample(1:5000, 4, 0),]; print(paste(nrow(h_net), "Observations"))
h_com <- read.csv(file = "./edges.csv")
h_com[sample(1:5000, 4, 0),]; print(paste(nrow(h_com), "Observations"))
```


## Data Cleaning

Before anything can be done with the data, it is necessary to clean it, the weight ascribed to an edge will be the number of interactions between any two characters, this is because it represents a measur of the strength of the relationship between two characters.

The easiest way to acheive this is to create an adjacency matrix wher ethe value of each cell is the weight of the edge and then melt that adjacency matrix into a long-format table:

```{r}
# Create Adjacency Matrix -------------------------------------------------
make_adj <- function(mat){
ig_adj <- get.adjacency(graph.edgelist(as.matrix(mat), directed=FALSE))
as.matrix(ig_adj)
}

# Plot Matrix -------------------------------------------------------------

# Rotate matrix 90-degrees clockwise.
rotate <- function(x) {
    t(apply(x, 2, rev))
}

# create the image
plot_mat <- function(mat){
image(rotate(mat), col = c(3, 6), axes = FALSE, frame.plot = TRUE, main = "Values > 10")
}

data_adj <- make_adj(h_net)
if(!((colnames(data_adj)==rownames(data_adj)) %>% mean())) print("The Adjacency Matrix is not symmetrical about i=j")
mean_social <- apply(data_adj, 1, mean)
data <- melt(data_adj)
data <- cbind(data, mean_social)
data <- data[data$value > 0,]
data <- data[order(-data$mean_social),] # This is the wrong order, I should use mean level
names(data) <- c("Source", "Target", "Value", "Mean_Social")
data_all <- data
head(data_all)
```

### Sample the Data

One of the issues with this data set however is that there is simply too much data to visualise effecively (or frankly to compute in any reasonable time frane in a scripting language) and so instead the data will be subset:

```{r}
dsamp_social <- function(n, seed) {
index <- sample(1:nrow(data), size = n, replace = FALSE)
data_all[index,]
}

data <- dsamp_social(990, 309320932932)
```


## Relational Investigation

Primarily this dataset is relational without direction or heirarchy and so the most appropriate method to visualise it would be with a graph depicting the inter-relation of characters via edges and vertices.

### 2D Network Graph

A 2d Network Graph describing the inter-relations of vertices and edges can be built using the `igraph` library:

```{r}
Edges <- data
G <- igraph::graph_from_data_frame(d = Edges, directed = FALSE)
   laytg <- igraph::layout.auto(G)
   colnames(laytg) <- c("xval", "yval")
   laytg <- as_tibble(laytg)
   laytg$node <- vertex_attr(G)[[1]]
   laytg %>% head()
```

With the vertices described it is necessary to build the lines by creating a data frame of the starting and Ending points of edges:

```{r}
ne <- nrow(Edges)
   ys <- xe <- ye <- xs <- vector(length = ne)
   for (i in seq_len(ne)) {
     xs[i] <- laytg$xval[laytg$node==Edges$Source[i]]
     ys[i] <- laytg$yval[laytg$node==Edges$Source[i]]
     xe[i] <- laytg$xval[laytg$node==Edges$Target[i]]
     ye[i] <- laytg$yval[laytg$node==Edges$Target[i]]
   }
  
   EdgeNames <- paste0("From:", Edges$Source, "To:", Edges$Target) 
   starts    <- data.frame("xval" = xs, "yval" = ys, edgeid = EdgeNames) # TODO Make a factor
   ends      <- data.frame("xval" = xe, "yval" = ye, edgeid = EdgeNames)
   Edges_val <- as_tibble(rbind(starts,ends))
   Edges_val %>% tail()
```

Using `ggplot2` the vertices can be plotted as points and the
edges as lines, it is important to specify that the group the lines belong to
are the names of the edges (`edgeid` values) because otherwise all the lines
will be connected, this may be acceptable for a graph where all vertices are
connected by edges, but for this data it is quite possible in this data set that two characters
never interact (for example the *Avengers* might not interact with the *X-men*
or the Gods of the Aesir might not interact with the *Fantastic Four* etc.), 
such a graph is known as a **_disconnected_** *graph* [@weisstein].


```{r}
  p <-  ggplot(rbind(starts,ends), aes(x = xval, y = yval)) +
        geom_line(aes(group = edgeid)) +
        geom_point(data = laytg, aes(x = xval, y = yval, col = node), size = 2) +
        theme_classic()
  ggplotly(p)
```

Unfourtunately this method produces a lot of islands and the interactions are difficult to understand because of the edge overlaps, so instead of taking a sample, perhaps it would be better instead to choose from the characters that have the highest absolute cross overs rather than the highest average, this will make the data less spread out:

```{r}
data_all <- data_all[order(-data_all$Value),] # This is the wrong order, I should use mean level
data <- data_all[1:300,]
```

And so recreating the graph:

```{r}
Edges <- data
gg_network_graph <- function() {
G <- igraph::graph_from_data_frame(d = Edges, directed = FALSE)
   laytg <- igraph::layout.auto(G)
   colnames(laytg) <- c("xval", "yval")
   laytg <- as_tibble(laytg)
   laytg$node <- vertex_attr(G)[[1]]
   laytg %>% head()
   
   
ne <- nrow(Edges)
   ys <- xe <- ye <- xs <- vector(length = ne)
   for (i in seq_len(ne)) {
     xs[i] <- laytg$xval[laytg$node==Edges$Source[i]]
     ys[i] <- laytg$yval[laytg$node==Edges$Source[i]]
     xe[i] <- laytg$xval[laytg$node==Edges$Target[i]]
     ye[i] <- laytg$yval[laytg$node==Edges$Target[i]]
   }
  
   EdgeNames <- paste0("From:", Edges$Source, "To:", Edges$Target) 
   starts    <- data.frame("xval" = xs, "yval" = ys, edgeid = EdgeNames) # TODO Make a factor
   ends      <- data.frame("xval" = xe, "yval" = ye, edgeid = EdgeNames)
   Edges_val <- as_tibble(rbind(starts,ends))
   Edges_val %>% tail()
   
   
  p <-  ggplot(rbind(starts,ends), aes(x = xval, y = yval)) +
        geom_line(aes(group = edgeid), size = 0.1, lty = 3) +
        geom_point(data = laytg, aes(x = xval, y = yval, col = node), size = 2.5) +
        labs(x = "", y = "") +
        theme_classic() +
        theme(axis.line = element_blank(),  # https://stackoverflow.com/a/6542792/12843551
              axis.text.y=element_blank(),axis.ticks=element_blank(),
              axis.text.x=element_blank())
  return(p)
}

ggplotly(gg_network_graph())

```

Although this method of sampling the data is less of a fair sample (biasing selection for characters that often interact such as the *X-Men*, *Fantastic Four* and *Avengers*) it does provide for an effective way to observe the most frequent form of interaction between any two characters as opposed to trying to see every interaction between all characters through a random sample; for this reason this form of sampling is more appropriate for making a visualisation of the relationships between characters and will be used henceforth.



#### Visualisation Techniques

For this plot 300 edges were chosen because the interactive nature of the plot is such that the overlap of edges can be made arbitrarily small by zooming sufficiently far in, given that there are distinct islands this is necessary in order to visualise the data and hence more data points may as well be visualised (Select an area with the mouse to zoom in and right click to zoom out).

It was necessary to balance the need to zoom in with the size of the vertices/points because the points will respond to a mouse hover by returning the name of the node.

Genereally only a small amount of colours can be used to distinguish variables in a plot, however in this case the variables are going to be considered relative to other local vertices and so multiple colours can be an effective way to distinguish nodes in relative proximity to one another.

Although the legend is quite long and not useful on it's own, in conjunction with the capacity to adjust the range of the plot by zooming and the fact that the legend can be scrolled it becomes an effective way to compare multiple plots that are in close proximity to each other as a reference second to the mouse-hover dialog.


#### Very Large Sample

A much larger sample can be used to get an understanding of how the often seperate disconnected 'islands' of characters might become connected by another character taken into the snapshot:

```{r}
Edges <- data_all[1:600,]
# Edges <- data_all[1:999,]
ggplotly(gg_network_graph())

```

Although such a large graph is inherently difficult to interpret for clusters close to one another it is useful for observing broader relations in conjuction with the zooming, panning and using the mouse-hover feature.

#### Smaller Sample

A graph destined for physical print will be more appropriately constructed if the scope is restricted to a smaller number of characters and annotations are added to the vertices such that the audience can come to understand the potential connections between the characters (as opposed to a broad overview that can be adjusted in digital media):

```{r}
gg_network_graph_print <- function() {
G <- igraph::graph_from_data_frame(d = Edges, directed = FALSE)
   laytg <- igraph::layout.auto(G)
   colnames(laytg) <- c("xval", "yval")
   laytg <- as_tibble(laytg)
   laytg$node <- vertex_attr(G)[[1]]
   laytg %>% head()
   
   
ne <- nrow(Edges)
   ys <- xe <- ye <- xs <- vector(length = ne)
   for (i in seq_len(ne)) {
     xs[i] <- laytg$xval[laytg$node==Edges$Source[i]]
     ys[i] <- laytg$yval[laytg$node==Edges$Source[i]]
     xe[i] <- laytg$xval[laytg$node==Edges$Target[i]]
     ye[i] <- laytg$yval[laytg$node==Edges$Target[i]]
   }
  
   EdgeNames <- paste0("From:", Edges$Source, "To:", Edges$Target) 
   starts    <- data.frame("xval" = xs, "yval" = ys, edgeid = EdgeNames) # TODO Make a factor
   ends      <- data.frame("xval" = xe, "yval" = ye, edgeid = EdgeNames)
   Edges_val <- as_tibble(rbind(starts,ends))
   Edges_val %>% tail()
   
   
  p <-  ggplot(rbind(starts,ends), aes(x = xval, y = yval)) +
        geom_line(aes(group = edgeid), size = 0.6, lty = 3) +
        geom_point(data = laytg, aes(x = xval, y = yval, col = node), size = 2.5) +
        geom_label(data = laytg, aes(x = xval, y = yval, label = node), size = 1.5, nudge_x = 0.3, nudge_y = 0.8, ) +
        labs(x = "", y = "") +
        guides(col = FALSE) +
        theme_classic() +
        theme(axis.line = element_blank(),  # https://stackoverflow.com/a/6542792/12843551
              axis.text.y=element_blank(),axis.ticks=element_blank(),
              axis.text.x=element_blank())
  return(p)
}


Edges <- data_all[1:40,]
gg_network_graph_print()

```

### 3D Network Graph

Although a 2D graph can provide a better seperation of nodes the capacity to visualise more elements by way of a third dimension can provide a deeper understanding of the connections between the data.

In order to create a 3d graph the [*Plotly*](https://plotly.com/) graphing library may be used, this is built on the [D3JS](https://d3js.org/) library and geared towards data science and AI.


#### Smaller Sample

##### Build the Graph

First Subset the Data and greate an `igraph` object:

```{r}
Edges <- data_all[1:300,]

# Get the number of nodes and edges
nodes <- factor(c(as.character(data$Source), as.character(data$Target)))
nodes <- nodes[!duplicated(nodes)]
nn <- length(nodes)
ne <- nrow(data)

# Build the Graph
G <- igraph::graph_from_data_frame(Edges, directed = FALSE)

```

##### Layout the Nodes
From the constructed `igraph` object a layout of nodes in 3 dimensions may be produced:

```{r}
# Get the node positions set by the layout for 3D graphs (I like Kamada Kawai)
layt <- as.data.frame(igraph::layout.kamada.kawai(G, dim = 3))
names(layt) <- c("xval", "yval", "zval")
layt$label <- vertex_attr(G)[[1]]

# Get the X Co-ordinates of the Nodes
   xn <- layt[,1] # X co-ordinate of nodes
   yn <- layt[,2]
   zn <- layt[,3]
```



##### Replace the Edges with the Node Values

The corresponding endpoints for the edges can be taken from the listed nodes:

```{r}
   Zs <- Ys <- Xs <- Ze <- Ye <- Xe <- vector(length = ne)
   for (i in seq_len(ne)) {
     Xs[i] <- layt$xval[layt$label==Edges$Source[i]]
     Ys[i] <- layt$yval[layt$label==Edges$Source[i]]
     Zs[i] <- layt$zval[layt$label==Edges$Source[i]]
     Xe[i] <- layt$xval[layt$label==Edges$Target[i]]
     Ye[i] <- layt$yval[layt$label==Edges$Target[i]]
     Ze[i] <- layt$zval[layt$label==Edges$Target[i]]
   }
   
   starts    <- data.frame("xval" = Xs, "yval" = Ys, "zval" = Zs, edgenum = 1:ne) # TODO Make a factor
   ends      <- data.frame("xval" = Xe, "yval" = Ye, "zval" = Ze, edgenum = 1:ne)
   Edges_val <- as_tibble(rbind(starts,ends))
   Edges_val
```

##### Create the Plotly Elements

###### Create the Points Vertices

The points can then be built with plotly:
```{r}
ponts <- plot_ly(x = layt[,1],
                 y = layt[,2],
                 z = layt[,3],
                 text = layt$label,
                 color = layt$label,
                # text = 0:(nn-1),
                type = "scatter3d",
                mode = "markers")
```

###### Create the Lines (Edges)

Building the edges is a little tricker, the plotly doesn't support grouping line elements and so each edge must be produced as a seperate plot and then combined as submodules (if the graph was assured to be completely connected then the endpoints could simply be provided and all connected):

```{r, echo = FALSE, results = FALSE, quietly = TRUE, warning=FALSE}
line_plots <- list()   # Does not support groups so you must create a seperate line plot so it doesn't just join everything
for (i in 1:ne) {                # the idea is to create a different sub plot for every line.
line_plots[i] <- plot_ly(
    x = c(starts$xval[i], ends$xval[i]),
    y = c(starts$yval[i], ends$yval[i]),
    z = c(starts$zval[i], ends$zval[i]),
    mode = 'lines',
    type = "scatter3d",
    showlegend = FALSE) # You don't want each edge to be labelled, only the nodes
}
lines_grouped <- subplot(line_plots)

```

###### Remove Axis Elements

The axis Elements are not desirable a network graph:

```{r}
# Pull out background.
noback <- list(showticks = FALSE,
               showgrid = FALSE,
               showaxis = FALSE,
               gridcolor = "white",
               showline=FALSE,
               zeroline=FALSE,
               showticklabels=FALSE, title='')


# lines <- lines %>% layout(scene = list(xaxis=noback,yaxis=noback,zaxis=noback))
lines_grouped <- lines_grouped %>% layout(scene = list(xaxis=noback,yaxis=noback,zaxis=noback))
ponts <- ponts %>% layout(scene = list(xaxis=noback,yaxis=noback,zaxis=noback)) # you need to do this for points to get rid of x,y,z labels
```

###### Put it all Together

All the points can now be combined to create the network graph:

* Hold Right-Mouse button to pan
* Hold Left-Mouse button to rotate orbitally
* Scroll in order to zoom in/out.

```r
 subplot(lines_grouped, ponts)
```

```{r, warning=FALSE, quietly = TRUE, echo = FALSE, message=FALSE}
suppressWarnings(subplot(lines_grouped, ponts))
```

Although the points are very large, after zooming in to a degree that the edges do not overalap too much, the size of the nodes is sufficiently large to respond to a mouse hover yet not too large to obstruct from the sahpe of the graph and so such a size was settled on.


#### Large Sample

This can be repeated for a much larger sample in order to get a more immersive (although less clear) visualisation of the data, this can be useful for observing large patterns but less useful for identifying individual relations:


```{r, warning=FALSE, quietly = TRUE, echo = FALSE, message=FALSE}
data <- data_all[1:150,] # previously 1000

# Get the number of nodes and edges
nodes <- factor(c(as.character(data$Source), as.character(data$Target)))
nodes <- nodes[!duplicated(nodes)]
nn <- length(nodes)
ne <- nrow(data)

# Build the Graph
Edges <- data
G <- igraph::graph_from_data_frame(Edges, directed = FALSE)

# Get the node positions set by the lKamada-Kawai layout for 3D graphs 
layt <- as.data.frame(igraph::layout.kamada.kawai(G, dim = 3))
names(layt) <- c("xval", "yval", "zval")
layt$label <- vertex_attr(G)[[1]]
head(layt)
   
## The order is not preserved
(factor(vertex_attr(G)[[1]])==factor(nodes)) %>% mean()


#   layt <- igraph::layout.auto(G, dim = 3)
   
   # Set data for the Plotly plot of the graph:
   
   xn <- layt[,1] # X co-ordinate of nodes
   yn <- layt[,2]
   zn <- layt[,3]
   
   Zs <- Ys <- Xs <- Ze <- Ye <- Xe <- vector(length = ne)
   for (i in seq_len(ne)) {
     Xs[i] <- layt$xval[layt$label==Edges$Source[i]]
     Ys[i] <- layt$yval[layt$label==Edges$Source[i]]
     Zs[i] <- layt$zval[layt$label==Edges$Source[i]]
     Xe[i] <- layt$xval[layt$label==Edges$Target[i]]
     Ye[i] <- layt$yval[layt$label==Edges$Target[i]]
     Ze[i] <- layt$zval[layt$label==Edges$Target[i]]
   }
   
   starts    <- data.frame("xval" = Xs, "yval" = Ys, "zval" = Zs, edgenum = 1:ne) # TODO Make a factor
   ends      <- data.frame("xval" = Xe, "yval" = Ye, "zval" = Ze, edgenum = 1:ne)
   Edges_val <- as_tibble(rbind(starts,ends))
   
   
   
   # Start Building the Plotly
   
   # Plotly
ponts <- plot_ly(x = layt[,1],
                 y = layt[,2],
                 z = layt[,3],
                 text = layt$label,
                 color = layt$label,
                 marker = list(size = 4),
                # text = 0:(nn-1),
                type = "scatter3d",
                mode = "markers")



line_plots <- list()   # Does not support groups so you must create a seperate line plot so it doesn't just join everything
for (i in 1:ne) {                # the idea is to create a different sub plot for every line.
line_plots[i] <- plot_ly(
    x = c(starts$xval[i], ends$xval[i]),
    y = c(starts$yval[i], ends$yval[i]),
    z = c(starts$zval[i], ends$zval[i]),
    type = "scatter3d",
    mode = 'lines',
    showlegend = FALSE) # You don't want each edge to be labelled, only the nodes
}
lines_grouped <- subplot(line_plots)

# https://plotly.com/python/v3/3d-network-graph/

# Pull out background.
noback <- list(showticks = FALSE,
               showgrid = FALSE,
               showaxis = FALSE,
               gridcolor = "white",
               showline=FALSE,
               zeroline=FALSE,
               showticklabels=FALSE, title='')


lines_grouped <- lines_grouped %>% layout(scene = list(xaxis=noback,yaxis=noback,zaxis=noback))
ponts <- ponts %>% layout(scene = list(xaxis=noback,yaxis=noback,zaxis=noback)) # you need to do this for points to get rid of x,y,z labels

# subplot(ponts, lines)
# subplot(lines_grouped, ponts) %>% suppressWarnings() %>% suppressMessages()
   
```

Although the large vertex size does obstruct the fuzzy shape created by the edges, they were elected for as a compromise for the benefit of the mouse-hover annotations that allow a character to be tied to a node.

This graph has too many nodes to show relations between individuals but can be useful to show relations between clusters, for example the Gods of the Aesir/Vanir appear as a distinct cluster linked by Thor and and Captain American can be seen to be a very sentral character that acts as the main tie between much of the universe.

### Results

The relational network graphs show generally that the Marvel Characters tend to form there own clusters, as more characters are added (in order of the number of cross overs that character has had generally), these clusters become more and more connected through different characcters.

The large network graph shows two very interesting features, primarly that certain characters such the Gods are a cluster quite far from most other characters, representing the fact that Gods don't interact with humans often in the comics. Secondly it shows that characters such as *Iront Man*, *Captain America* and *Thor* are very central to all other characters, having many cross-overs generally with a wide-diversity of characters, demonstrating that these core-characters represent a central binding tool to tie the diverse [^1] stories together into one universe.

[^1]: Diverse being an understatement given that the characters contained in this data set range from Jack Frost to Odin to Adolf Hitler to Hercules with settings ranging from the Second World War to the Ghettos of US Cities.
Islands are cool, captain america connects everything





