---
title: "02 Quizz 1; Chi Dist"
output: 
  html_document: 
    keep_md: yes
editor_options: 
  chunk_output_type: inline
---

# Quizz 1

## Q1; Comparison of Observation to Base Ratio

A set of 104 university students were surveyed, asking for their favourite pizza type, with the following results.

| Hawaiian | 	Barbecue | 	Supreme |
| --- | 	--- | 	--- |
| 32 | 	14 | 	58 |

The Australian population has the following pizza preferences.

| Hawaian | Barbecue | Supreme |
| ---     | ---      | ---     |
| 0.5     | 0.3      | 0.2     |

Calculate the chi squared distance between the student sample and the Australian population.

### Solution

First enter the values, really carefully:

```{r}
pizza <- c(32, 14, 58)
aus <- c(0.5, .3, 0.2)
```


This can either be calculated manually:

```{r}
e <- sum(pizza)*aus
o <- pizza

(chival <- sum((e-o)^2/e))
```

Or the built in function can be used:

```{r}
chisq.test(pizza, p = aus, rescale.p = TRUE)
```


## Q2; Comparing Populations (Assuming they're identically distributed)

A survery was performed on 100 people, asking for their city of birth and their
eye colour. The results are tabulated below.

| Brown     | Hazel |  Blue |       |
| ---       |   --- |   --- |   --- |
| Sydney    | 10.00 | 30.00 | 20.00 |
| Melbourne | 20.00 | 10.00 | 10.00 |

If we assume that city of birth has no effect on eye colour, calculate the
expected number of people that are born in Melbourne and have brown eyes.

### Solution

So basically here:

* H0
  + Populations are distributed with equivalent Proportions
* Ha  
  + Populations have different Proportions.
  
### First Enter the Data
  
```{r}
# Create Vectors
sydney    <- c("Brown" = 10,"Hazel" = 30,"Blue" = 20)
melbourne <- c("Brown" = 20,"Hazel" = 10,"Blue" = 10)
```

  
#### Create a Matrix
```{r}
eye_Mat <- rbind(sydney, melbourne)
```

### Determine the expected distributions

#### Outer Product Method

```{r}

obs_count <-  rowSums(eye_Mat)
feature_proportions  <- colSums(eye_Mat)/sum(eye_Mat)

outer(rowSums(eye_Mat), colSums(eye_Mat)/sum(eye_Mat))

```

#### Matrix Method

```{r}
obs_count
feature_proportions 
as.matrix(obs_count) %*% t(as.matrix(feature_proportions)) 
```

#### Row and Col Sum Method

Another way to remember it is that the expected value is:

$$
e_{ij} = \frac{\textsf{rowTotal} * \textsf{colTotal}}{\textsf{grandTotal}}
$$


### Perform a Chi Test

```{r}
(cht <- chisq.test(x = eye_Mat, simulate.p.value = TRUE, B = 10^4))
```

Hence the probability of there detecting a difference between the two populations, under the assumption that the populations are identical is `r round(cht$p.value, 3)`


## Q3; Test For Uniform Distribution

A Bank is open everyday from 10am to 3pm and closed for lunch between 12pm and
1am. For one particular day the tellers serve the following number of customers
in each hour.

| 10-11 | 11-12 |   1-2 |   2-3 |
|   --- |   --- |   --- |   --- |
| 80.00 | 70.00 | 69.00 | 61.00 |

Management would like to know if this is consistent with a uniform distribution
of customers across the day. What is the expected number of customers per hour,
if a uniform distribution is correct?

### Solution

A uniform distribution would be such that each value is equal, hence:

```{r}
bank <- c(80, 70, 69, 61)
e <- mean(bank)
```

The probability of incorrectly asserting that the values do differ over time (under the assumption that all values are identical) is given by:

```{r}
chisq.test(bank, rescale.p = TRUE)
```

This is equivalent to specifying equal proportions, that's the default assumption made by **_R_**:

```{r}
chisq.test(bank, p = c(0.25, 0.25, 0.25, 0.25), rescale.p = TRUE)
```

In this case there is not enough evidence to reject the assumption that the distriution varies with time.


## Q4; Expected Frequencies

A large sample survey of Australian businesses showed the following percentage
of computers running the following operating systems.

| Windows | OS X  | Linux |
| ---     | ---   | ---   |
| 30.00   | 50.00 | 20.00 |


Compute the expected number of computers using Linux from a sample of size 32.

### Solution

Just multiply the proportion through:

```{r}
comp <- c(30, 50, 20)
comp_rate <- comp/sum(comp)

comp_rate[3]*32
```

hence the expected number of computers running linux is 6.4.




