---
title: "01 Randomness (Pi Digits)"
tags: [Notebooks/Thinking About Data]
output: html_notebook
---


# (01) Randomness

## Preamble

```{R}
load.pac <- function() {
  
  if(require("pacman")){
    library(pacman)
  }else{
    install.packages("pacman")
    library(pacman)
  }
  
  pacman::p_load(xts, sp, gstat, ggplot2, rmarkdown, reshape2, ggmap,
                 parallel, dplyr, plotly, tidyverse, reticulate, UsingR, Rmpfr)
  
}

load.pac()
 
```

```{r setup, include=FALSE}
library(reticulate)
use_python("/usr//bin/python3")
```



## (1) Compute tables of $\pi$ digits
First it is necessary to have some digits of pi, we'll just take a dataset:

```{r}
pi_Dig <- UsingR::pi2000[-1]
write.csv(pi_Dig, file = "./piDigits.csv", quote = FALSE, row.names = FALSE)
pi_Tibble <- tibble::enframe(piDig)
tibble::remove_rownames(pi_Tibble)
pi_Tibble <- pi_Tibble$value
head(pi_Tibble)
```
### using the `Rmpfr` package

An alternative is to use the Rmpfr package deals with handling numbers of arbitrary precision, this requires the `GMP C` library to be installed, so the package will be something like `libbmp-dev` and should be in the repos for `apt` or `pacman -S` or whatever:

```{bash}
apt list libgmp3-dev
apt list libgmp
apt list libmpfr-dev


apt list libgmp3-dev libgmp libmpfr-dev

```


Rather than specifying significant figures the package works with bits of precision, so for example a base 10 number like 8 can be represented with 3 bits of information because $2^3=8$.

if 1000 decimal places were required, the number of bits would be the number of binary values necessary to represent that same value:

$$\begin{aligned}
2^{\textsf{bits}} &= 10^{\textsf{digits}} \\
\iff \textsf{bits} &= \textsf{digits} \times \log_{2}{}{\left(  10\right) }\\
\end{aligned}$$


```{r}
library(Rmpfr)
diglength <- 2000 

precision <- diglength*log2(10)
precision <- ceiling(precision)
piVal <- Rmpfr::Const("pi", precision)
print(Rmpfr::Const("pi", 12*log2(10)))
```

In order to extract the value use `substring()` in order to create substrings of the values. 

```{r}
piVal      <- format(piVal)
pi_Digits  <- substring(text = piVal, first = 1:diglength, last = 1:diglength)[3:diglength] #%>% as.numeric()
pi_Digits  <- substring(text = piVal, first = 1:diglength, last = 1:diglength)[3:diglength] %>% as.integer()
pi_Digits <- as.vector(pi_Digits)
#pi_Digits <- factor(pi_Digits, levels = 0:9, ordered = TRUE)
```

A histogram of which may be generated:

```{r}
table(pi_Digits)
table(pi_Digits) %>% barplot
```


The problem with a histogram is that it will combine The first two frequencies in a way that is incorrect in this context, it only seems to do it with the pi digits though for some reason

```{r}
hist(pi_Digits, breaks = seq(from = 0, to = 9, by = 1)) %>% summary()
axis(side = 1, at = seq(0, 9, 1), labels = seq(0, 9, 1))
axis(2,pos = -9)


random_Uniform <- runif(2000, 0, 9)
hist(random_Uniform)
axis(side = 1, at = seq(0, 9, 1), labels = seq(0, 9, 1))
axis(2,pos = -9)
```

A better alternative is to use ggplot2

```{r}
pi_DigitsDF <- tibble::enframe(pi_Digits)

HistPlot <- function (DataFrame, heading) {
ggplot(data = DataFrame, aes(x = value, fill = value)) +
#       geom_histogram(binwidth = 1, fill = "lightblue", col = "red") +
       geom_histogram(bins = 10, fill = "lightblue", col = "red") +
       theme_classic() +
       scale_x_continuous(breaks = seq(0, 10, 1)) +
       labs(x = "Decimal Value", y = "Frequency",
	    title = heading)
    
}

HistPlot(pi_DigitsDF, "Distribution of Digits of Pi" )
```




### Using Power Series

Using integration by parts, it can be shown that:

$$\begin{aligned}
\int^{b}_{0} \frac{1}{1+ x^2}  \mathrm{d}x&= \arctan\left( b \right) \enspace : \quad b \in\left( \mathbb{R}^+\cap 0 \right)  \\
 \implies  \int^{1}_{0} \frac{1}{1+ x^2}  \mathrm{d}x&= \arctan\left( 1 \right) \\
 \implies  \int^{1}_{0} \frac{1}{1+ x^2}  \mathrm{d}x&= \arctan\left( 1 \right) \\
\end{aligned}$$
Transforming into a a power series with some other magic:

$$\begin{aligned}
\implies  \pi &= 4 \times \sum^\infty_{n=0}\left[ \frac{\left( - 1 \right) ^k}{2k+ 1}   \right] 
\end{aligned}$$

The problem with this is that the convergence is too slow to be useful:

$$\begin{aligned}
\frac{1}{2k+ 1} \leq 10^{100}  \implies  k\geq 10^{100} \times \frac{1}{2} - 1
\end{aligned}$$

presuming 1 sum per cycle at 5 GHz this would take well in excess of the age of the universe.
There are ways to speed it up but it's all well outside scope





## (4) Evaluate Distribution of pi Digits

In order to evaluate whether or not the digits of pi appear uniformally distriuted first consider the first 50 digits:

### First 50 Digits of pi

```{r}
count <- table(pi_Digits[1:50])
rss   <- (table(pi_Digits[1:50])-50/10)^2

piDigErrorDF        <- data.frame(0:9, as.vector(count), as.vector(rss))
names(piDigErrorDF) <- c("value", "Count", "SquareError")

print(piDigErrorDF)
SSE <- sum(piDigErrorDF$SquareError)

print(paste("The Sum of Squared Errors is", SSE))
HistPlot(pi_DigitsDF[1:50,], "Distribution of first 50 Digits of Pi")
```

### First 200 Digits of Pi

```{r}
count <- table(pi_Digits[1:200])
rss   <- (table(pi_Digits[1:200])-200/10)^2

piDigErrorDF        <- data.frame(0:9, as.vector(count), as.vector(rss))
names(piDigErrorDF) <- c("value", "Count", "SquareError")

print(piDigErrorDF)
SSE <- sum(piDigErrorDF$SquareError)

print(paste("The Sum of Squared Errors is", SSE))
HistPlot(pi_DigitsDF[1:200,], "Distribution of first 200 Digits of Pi")
```

### First 500 Digits of Pi

```{r}
count <- table(pi_Digits[1:500])
rss   <- (table(pi_Digits[1:500])-500/10)^2

piDigErrorDF        <- data.frame(0:9, as.vector(count), as.vector(rss))
names(piDigErrorDF) <- c("value", "Count", "SquareError")

print(piDigErrorDF)
SSE <- sum(piDigErrorDF$SquareError)

print(paste("The Sum of Squared Errors is", SSE))
HistPlot(pi_DigitsDF[1:500,], "Distribution of first 500 Digits of Pi")
```


### Random Sample of pi Digits

In **_R_** sampling something with repetition is referred to replacing, so to count or sample something with repetition specify `replace = TRUE`.
 Recall the counting Formulas:

 | selection | ordered | unordered |
 | --- | --- | --- |
 | With Repetition | $n^m$ | $\binom{m+n-1}{n}$ |
 | Without Repetition | $n_{(m)}$ | $\binom{n}{m}$ |
 
 Where:
 
 * $\binom{n}{m} =\frac{m_(k)}{k!}=\frac{m!}{k!(m-k)!}$
 * $n_{(m)}=\frac{n!}{(n-m)!}$
 * $n! = n \times (n-1) \times (n-2) \times 2 \times 1$

In order to randomly sample the digits of Pi:

```{r}
index <- sample(1:diglength, size = 200)
print(table(pi_Digits[index]))
HistPlot(pi_DigitsDF[index,], "Distribution of first 50 Digits of Pi")
```

This can be repeated multiple times:

```{r}
library(gridExtra)
library(tidyverse)

PlotList <- list()
for (i in 1:6) {
index <- sample(1:diglength, size = 200)
PlotList[[i]] <- HistPlot(pi_DigitsDF[index,],
			  paste("Distribution of a random sample of \n 200 Digits of Pi from the first", diglength))
}

# arrangeGrob(grobs = PlotList, layout_matrix = matrix(1:6, nrow = 3))
grid.arrange(grobs = PlotList, layout_matrix = matrix(1:6, nrow = 3))

```


## (5) Uniformally Distributed Values

### 50 Digits

```{r}
x <- runif(50, 0, 9) %>% round()
x <- sample(0:9, replace = TRUE, size = 50)
xDF <- tibble::enframe(x)

count <- table(x[1:50])
rss   <- (table(x[1:50])-50/10)^2

ErrorDF        <- data.frame(0:9, as.vector(count), as.vector(rss))
names(ErrorDF) <- c("value", "Count", "SquareError")

print(piDigErrorDF)
SSE <- sum(piDigErrorDF$SquareError)

print(paste("The Sum of Squared Errors is", SSE))
HistPlot(xDF[1:50,], "Distribution of a uniform random sample of 50 digits")
```

### 200 Digits


```{r}
x <- runif(200, 0, 9) %>% round()
x <- sample(0:9, replace = TRUE, size = 200)
xDF <- tibble::enframe(x)

count <- table(x[1:200])
rss   <- (table(x[1:200])-200/10)^2

ErrorDF        <- data.frame(0:9, as.vector(count), as.vector(rss))
names(ErrorDF) <- c("value", "Count", "SquareError")

print(piDigErrorDF)
SSE <- sum(piDigErrorDF$SquareError)

print(paste("The Sum of Squared Errors is", SSE))
HistPlot(xDF[1:200,], "Distribution of a uniform random sample of 200 digits")
```

### 500 Digits


```{r}
x <- runif(500, 0, 9) %>% round()
x <- sample(0:9, replace = TRUE, size = 500)
xDF <- tibble::enframe(x)

count <- table(x[1:500])
rss   <- (table(x[1:500])-500/10)^2

ErrorDF        <- data.frame(0:9, as.vector(count), as.vector(rss))
names(ErrorDF) <- c("value", "Count", "SquareError")

print(piDigErrorDF)
SSE <- sum(piDigErrorDF$SquareError)

print(paste("The Sum of Squared Errors is", SSE))
HistPlot(xDF[1:500,], "Distribution of a uniform random sample of 500 digits")
```


## (6) Repeat for multiple larger digits
I did 50, 200 and 500 for all of them.

## (7) Pairs of Digits
Let's consider pairs of digits and their distribution:

```{r}
library(Rmpfr)
diglength <- 2000*2*10

precision <- diglength*log2(10)
precision <- ceiling(precision)
piVal <- Rmpfr::Const("pi", precision)
print(Rmpfr::Const("pi", 12*log2(10)))
```

In order to extract the value use `substring()` in order to create substrings of the values. 

```{r}
piVal      <- format(piVal); class(piVal)
pi_Digits  <- substring(text = piVal, first = seq(from = 1, to = diglength, by = 2), last = seq(from = 2, to = (diglength-1), by = 2))[3:(diglength/2)-1] %>% as.numeric()
pi_Digits  <- as.vector(pi_Digits)
#pi_Digits <- factor(pi_Digits, levels = 0:9, ordered = TRUE)
```

A histogram of which may be generated:

```{r}
table(pi_Digits)
table(pi_Digits) %>% barplot
```


A better alternative is to use ggplot2, count the bins carefully, 00 is it's own bin and so we would expect 99+1 bins overall

```{r}
pi_DigitsDF <- tibble::enframe(pi_Digits)

HistPlot <- function (DataFrame, heading) {
ggplot(data = DataFrame, aes(x = value, fill = value)) +
#       geom_histogram(binwidth = 1, fill = "lightblue", col = "red") +
       geom_histogram(bins = 100, fill = "lightblue", col = "red") +
       theme_classic() +
       scale_x_continuous(breaks = seq(0, 10, 1)) +
       labs(x = "Decimal Value", y = "Frequency",
	    title = heading)
    
}

HistPlot(pi_DigitsDF, "Distribution of Digits of Pi" )
```

This distribution looks mostly uniform, let's push it by doing a significantly larger analysis of pi:

```{r}
library(Rmpfr)
diglength <- 2000*2*10*10

precision <- diglength*log2(10)
precision <- ceiling(precision)
piVal <- Rmpfr::Const("pi", precision)
print(Rmpfr::Const("pi", 12*log2(10)))
```

In order to extract the value use `substring()` in order to create substrings of the values. 

```{r}
piVal      <- format(piVal); class(piVal)
pi_Digits  <- substring(text = piVal, first = seq(from = 1, to = diglength, by = 2), last = seq(from = 2, to = (diglength-1), by = 2))[3:(diglength/2)-1] %>% as.numeric()
pi_Digits  <- as.vector(pi_Digits)
#pi_Digits <- factor(pi_Digits, levels = 0:9, ordered = TRUE)
```

A histogram of which may be generated:

```{r}
table(pi_Digits)
table(pi_Digits) %>% barplot
```


A better alternative is to use ggplot2, count the bins carefully, 00 is it's own bin and so we would expect 99+1 bins overall

```{r}
pi_DigitsDF <- tibble::enframe(pi_Digits)

HistPlot <- function (DataFrame, heading) {
ggplot(data = DataFrame, aes(x = value, fill = value)) +
#       geom_histogram(binwidth = 1, fill = "lightblue", col = "red") +
       geom_histogram(bins = 100, fill = "lightblue", col = "red") +
       theme_classic() +
       scale_x_continuous(breaks = seq(0, 10, 1)) +
       labs(x = "Decimal Value", y = "Frequency",
	    title = heading)
    
}

HistPlot(pi_DigitsDF, "Distribution of Digits of Pi" )
```

