---
title: "08 Linear Regression"
fig_caption: yes
output:
  html_document: 
    keep_md: yes
    toc: yes
  html_notebook:
    toc: yes
  pdf_document:
    toc: yes
tags: Notebooks/Thinking About Data
editor_options:
  chunk_output_type: inline
---

# (08) Linear Regression
## Preamble

```{r}
## Preamble

## Install Pacman
load.pac <- function() {
  
  if(require("pacman")){
    library(pacman)
  }else{
    install.packages("pacman")
    library(pacman)
  }
  
    pacman::p_load(xts, sp, gstat, ggplot2, rmarkdown, reshape2, ggmap,
                 parallel, dplyr, plotly, tidyverse, reticulate, UsingR, Rmpfr,
                 swirl, corrplot, gridExtra, mise, latex2exp, tree, rpart, lattice,
                 coin, primes, epitools, maps, clipr, ggmap, RColorBrewer)
  
  
   mise()
   select <- dplyr::select
}

load.pac()
setwd(dir = "/home/ryan/Notes/DataSci/ThinkingAboutData/")
load(file = "~/Notes/DataSci/ThinkingAboutData/TAD.rdata")
load(file = "./TAD.rdata")
print("Success")
```


```{r}
knitr::opts_chunk$set(fig.path = "./figure/")

```

## Scatter Plot
A scatter plot can be created using the `plot` and `abline()` functions:

```{r, warning = FALSE}
head(crabsmolt)
plot(postsz ~ presz, data = crabsmolt)
abline(a = 0, b = 1)
abline(a = 18,  b = 1)
```

This could also be constructed with ggplot2:

```{r, warning = FALSE}
ggplot(crabsmolt, aes(x = presz, y = postsz)) +
    geom_point() +
    geom_smooth(method = lm, se = FALSE, col="blue") +
    geom_abline(slope = 1, intercept = 18, col = "red") +
    theme_bw()
```


## Linear Regression
### Crabs
```{r, warning = FALSE}
(fit = lm(postsz ~ presz, data = crabsmolt))formula

plot(postsz ~ presz, data = crabsmolt)
abline(fit)
```
### Pulse

```{r, warning = FALSE}
head(PulseBA)
formula <- "After ~ Before"
(pl.lm <- lm(formula, data = PulseBA))

plot(After ~ Before, data = PulseBA)
abline(pl.lm)
```

### Heights

```{r, warning = FALSE}
head(heights)
h.lm <- lm(earn ~ height, data = heights)

plot(y = heights$earn, x = heights$height)
abline(h.lm)
```

### Helmets

```{r, warning = FALSE}
head(nzhelmet)
nz.lm <- lm(Metal ~ Cardboard, data = nzhelmet)

plot(Metal ~ Cardboard, data = nzhelmet)
abline(nz.lm)
```


## Hypothesis Test (for no slope)
Given the hypothesis:

+ $H_0: \quad$ The slope has a value of 0
+ $H_a: \quad$ The slope has a non-zero value

In order to evaluate the probability of rejecting the null hypothesis when it is
true:

1. Simulate a sample corresponding to this population that has no relationship
  + This can be done by permuting the observations
2. Measure the number of times a slope value more extreme than observation is
recorded.
  + this represents the probability of a rejection corresponding to this sample
  being a /False Positive/.

```{r, warning = FALSE}
## Calculate the current slope
slope <- lm(postsz ~ presz, data = crabsmolt)[[1]][2]

## Simulate null hypothesis

sim <- replicate(10^3, {
    ## Shuffle the data
    postsz_perm <- sample(crabsmolt$postsz)
    ## Recalculate the slope
    slope_sim <- lm(postsz_perm ~ crabsmolt$presz, data = crabsmolt)[[1]][2]
    ## Is the slope more extreme
    abs(slope_sim) > abs(slope)
})

(pval <- mean(sim))
```

This simulation returns an expected p-value is 0 and hence the probability of rejecting
the null hypothesis when it is true, (i.e. the probability of rejecting the
hypothesis of no relationship when there actually is no relationship) is very
small and hence is rejected.

It is conceded that there is a relationship between the two variables at a very
high significance level.
  
### Pulse Data

The Hypothesis for the pulse data is:

+ $H_0: \quad$ There is no relationship between pulse rates before or after exercise
+ $H_a: \quad$ There is a non-zero rate of change for a model predicting
  Post-Exercise Pulse from Pre-Exercise.
  
The probability of rejecting the null hypothesis when it is correct can be
measured by simulating the frequency at which the null hypothesis would be
rejected if there was no relationship between the data in this sample given the
measured slope value.

```{r, warning = FALSE}
names(PulseBA)
## Calculate the Slope
slope <- lm(After ~ Before, data = PulseBA)[[1]][2]

## Simulate no Relationship
sim <- replicate(10^4, {
    ## Permutate the data so there is no relationship
    After_p <- sample(PulseBA$After)
    ## Measure the slope of a Linear Regression
    slope_sim <- coef(lm(After ~ Before, data = PulseBA))[2]
    ## Is this simulated value more extreme?
    abs(slope_sim) > abs(slope)
})

(pval <- mean(sim))

```

The expected p-value is returned as zero and so the probability of rejecting the
null hypothesis when it is true is very small, and hence the null hypothesis is
rejected.

It is conceded at a very high significance level that there is a non-zero rate
of change between the Post/Pre Pulse Rates in the population given this sample.

This doesn't provide that there is necessarily a high probability of correctly
rejecting the null hypothesis when it is false, this is measured by the power
$1 - \beta \approx \1-\textsf{FNR}$




### Exemplar Crabs
To test the hypothesis that the slope has some value

$$\begin{aligned}
y&= mx + c \\
y-mx& = c \\
y-mx& =  0\cdot m_2  + c \\
\end{aligned}$$

Now it is merely necessary to perform a hypothesis test for a zero slope for 
$y - mx \sim c$:

```{r, warning = FALSE}
## What is the observed slope
slope <- lm(postsz ~ presz, data = crabsmolt)[[1]][2]

## Simulate the observed slope when there is no relationship
sim <- replicate(10^4, {
    ## Permutate the Observations to remove a relationship
    presz_perm <- sample(crabsmolt$presz)
    ## Measure the slope
    slope_sim <- lm(postsz ~ presz_perm, data = crabsmolt)
    ## Is this slope more extreme?
    abs(slope_sim) > abs(slope)
})

## What is the frequency at which the null hypothesis would be rejected
    ## (erroneously if the observation was the cut off point)
(pval <- mean(sim))
```

The probability

```





## Confidence Interval (for slope)
## Residuals
## Summary Information
