---
title: "Housing Prices"
output: html_notebook
---

#Preamble
##Load Packages
First load the packages, the advantage to the `pacman` package is that it
will automatically install missing packages:

```{r}
if(require('pacman')){
  library('pacman')
}else{
  install.packages('pacman')
  library('pacman')
}

pacman::p_load(xts, ggplot2, rmarkdown, reshape2, ggmap,
               RColorBrewer, parallel, dplyr, plotly)

```
##Import Datasets
Now we need to import the datastets:

```{r}
house.df <- read.csv(file = "house2018.csv", header = TRUE, sep = ",")
head(house.df)
```

#Compare size and Price
If, on a fundamental level, the value of a property is dependent on its
size, this can be visualised with a plot:

```{r}
plot.house <- function() {
  
  plot(x = house.df$sqm, y = house.df$price,
                   type = "p", main = "House Prices",
                   xlab = "Property size (sqm)", ylab = "Price ($)",
                   frame.plot = TRUE, axes = TRUE,
                   col = "Indianred", pch = 16, cex = 2)
  }

plot.house()

```

##Create a linear model
A linear regression can be created for this relationship:

```{r}
nbhd      <- house.df$nbhd
offers    <- house.df$offers
sqm       <- house.df$sqm
brick     <- house.df$brick
bedrooms  <- house.df$bedrooms
bathrooms <- house.df$bathrooms
price     <- house.df$price

house.lm <- lm(price ~ sqm)
summary(house.lm)

```
Thus the linear model can be seen to be:
  $$Price =  7585 \times  Size - 101785$$
```{r}
plot.house.lm <- function(lm = TRUE) {
    plot.house()
  
  if(lm == TRUE) {
    abline(house.lm, col = "royalblue", lwd = 3, lty = 6)
  }
}

plot.house.lm(lm = TRUE)

```
 
 
Observe that the proportion of total variation explained by the linear model is only $r^2 = 30%$.

The large model residuals in this plot may be attributed as other variables
being misinterpreted as noise, hence we can use `ggplot2` to visualise the 
other data and more appropriately consider the bivariate relationship.

```{r}
ggplot(house.df, aes(x = sqm, y = price, shape = nbhd, label = brick, col = (bathrooms+ bedrooms))) +
  geom_point(size = 2) +
  guides(col = guide_legend(title = "Bathrooms + Bedrooms")) +
  guides(shape = guide_legend(title = "Neighbourhood")) +
  xlab("Size (sqm)") +
  ylab("Price $") +
  ggtitle("House Prices") + 
  geom_abline(slope = house.lm$coefficients[2],
              intercept = house.lm$coefficients[1], 
              lwd = 1, col = "indianred") +
  theme_classic()
```

#Multiple Linear Regression
Given that there are multiple factors influencing the price,
a multiple linear regression could present a potential model
for the plot:

```{r}
house.mlm <- lm(price ~ offers + sqm + bedrooms + bathrooms)
summary(house.mlm)
```

Thus a potential multiple linear regression could be:
$$Price = -174461 -136168 \times Offers + 6698 \times Size + 92571 \times bedrooms + 126113 \times bathrooms$$

#Answers
##1. Price a 210sqm house
The prediction for the price of a 210 sqm house would be:

```{r}
#always use assignments when using the predict() function
price <- house.df$price
sqm   <- house.df$sqm

predict(lm(price ~ sqm), data.frame(sqm=210), interval = "prediction", level = 0.95)

```

Using the Multiple Linear Regression would

```{r}



predict(house.mlm, data.frame(offers = mean(offers), sqm = 210, bedrooms = mean(bedrooms), bathrooms = mean(bathrooms)), interval = "prediction", level = 0.95)


```

Thus a 95% prediction interval for the value of a 210 sqm house would be centred around the bivariate model value of $\$1, 042, 669$ and range from:
$$\$1, 491, 120$ to $\$1, 939, 571$$.

##2. Value of a 15sqm extension
The modelled price of a property changes
in direct proportion to the size of the property
at a rate equivilant to the slope, thus 
a 15 sqm extension would be valued at:

```{r}
paste("$", round(house.lm$coefficients[2]) * 15)
```


#Appendix
Other relationships could also be readily visualised e.g. :

##Plot other variables with `ggplot2`
####Melt the dataframe
First the dataframe needs to be made tidy, 
the *nbhd* column is characters not numbers,
it could be made numerical using regex
and dplyr (or a loop) but that is outside the scope of
this work, the same can be said for the brick column,
these column will hence be omitted.
```{r}
house.melt <- melt(house.df[, -c(1:5)], id = c("price"))
head(house.melt, 3)
```

####Make the Plot
Now the data is tidy we can make the plot:

```{r}


ggplot(house.melt, aes(x = price, y = value, col = variable)) +
  geom_point() +
  xlab("Price ($)") +
  ylab("No. of Rooms") +
  ggtitle("Rooms and Housing Cost") + 
  theme_bw()


```




make a linear regression price~size sqm
  base plot
  ggplot2
make a multiple linear regression
  plot the regression with ggplot2??
