---
title: "Exercise 12.1"
output: 
  html_notebook: 
    toc: yes
---
The goal here is to create a prediction model for 
the value of benthic over the Chesepeake bay region.

This prediction model will be graphically portrayed with a:

1. Heatmap
    i) with base plot
    ii) wigh `ggplot2` over an actual map
2. Contour plot
    i) with base plot
    ii) With `ggplot2` over an actual map
3. 3d Surface plot
    i) with base plot
    ii) with a dynamic `plotly` surface plot


#Preamble
##Import Library

The following code will automatically install packages that are 
not already installed and load them into the library:

```{r}
if(require('pacman')){
  library('pacman')
}else{
  install.packages('pacman')
  library('pacman')
}

pacman::p_load(sp, EnvStats, gstat, ggplot2, rmarkdown, reshape2, ggmap, RColorBrewer, parallel, dplyr, plotly)
```

##Create Assignments

Note that  Benthic.df is included in the `EnvStats` package and is not a 
data frame but an `sp` class, which is very similar to a data frame

```{r}
#Assignments
benthic <- Benthic.df
lat   <- benthic$Latitude
lon   <- benthic$Longitude
index <- benthic$Index
```

#Inspect the Data Set

```{r}

head(benthic, 3)
```

#Lattitude and Longitude
* Lattitude is a line that runs east to west that measures how far 
north of the equatour a location is
      + Lattitude is assigned to the y-axis
* Longitude is a line that runs north to south that measures east/west
     + Longitude is assigned to the x-axis



#Assign the Co-ordinates
The *Benthic* Data set is an `sp` class objects and requires for its co-ordinates to be set:

```{r}
coordinates(benthic) = ~Longitude + Latitude
```
#Preliminary Step: Create a krige prediction Model
In order to create a krige prediction model two things are required:

* Fitted Variogram Model (e.g. an exponential model)
* A domain values over which to predict values

##Create a variogram Model




###Create a Variogram 
If a 4th degree polynomial model was used to model the 
benthic values over space, a variogram can be made from
that model thusly:

(The quadratic model is a required parameter of the kriging process,
in exercise 12.2 a quadratic model will be used,
the spacific mathematics of this though is somewhat beyond me)

```{r}
x <- lon
y <- lat

benthic.vg <- variogram(object = index ~ 
                          x + y + 
                          x^2 + x*y + y^2 + 
                          x^3 + x^2*y + x*y^2 + y^3 +
                          x^4 + x^3*y + x^2*y^2 + x*y^3 + y^4,
                        data = benthic)

head(benthic.vg)
```


###Fit an exponential model to that variogram

```{r}
benthic.vg.fit.exp <- fit.variogram(benthic.vg, model=vgm(1,"Exp", 0.5,1))
```

Thus the variogram model needed for the krige prediction has been ascertained

##Create a Grid of Domain values to use as predictor values
### Create a sequence of incrementally increasing for the X and Y axis
This will essentially act as the resolution of the overlay,
Most monitors are about 140 ppi, hence assuming the plot will be a
9 inch square, about 800 x 800 pixels should look right, but 
200 x 200 will be the used due to performance.
```{r}
xgrid       <-  seq(min(lon), max(lon), length.out = 200)
ygrid       <-  seq(min(lat), max(lat), length.out = 200)
```


###Combine into a domain base
This domain base however has to be a spatial object, this can be
acheived by defining co-ordinates in the data frame
```{r}
xy.surface  <- expand.grid(lon = xgrid, lat = ygrid)
coordinates(xy.surface) = ~lon + lat
head(xy.surface, 3)
```

##Create the Krige prediction object
```{r}

benthic.pred <- krige(formula = Index ~1,
                           locations =  benthic,
                           newdata = xy.surface,
                           model= benthic.vg.fit.exp)
head(benthic.pred)
```

Now there is a prediction model, if that prediction model is
extrapolated over our domain (`xy.surface`) the various plots can be
created

#Part 1: Create a heatmap

In order to create a heatmap, it is first necessary to take
the `sp` prediction object and make it into pixel data,
otherwise the heatmap won't look quite right:

```{r}
class(benthic.pred)
gridded(benthic.pred) = TRUE
class(benthic.pred)
```


```{r}
spplot(benthic.pred["var1.pred"],main="Predictions of Benthic Index - ordinary kriging")
```




#Part 2: Contour Map
In order to view the surface plot, contours can also be used.

 Recall from Week 11/Exercise 10, that contour and surface plots
  work a little differently to ordinary plots.
  
  A domain grid of equally spaced $x$ and $y$ values is required and
  a matrix of $z$ values that can be overlayed onto that domain grid
  in order to provide the surface values required 
  to create the surface plot.



  The method from Lecture 10/Wk. 11 involved predicting over
  the square domain of `xy.surface`, this won't work here
  because instead of the `predict` function, the `krige` function
  was used, hence it is necessary to convert the output from
  the `krige` function, into the output that would have been 
  given by the predict function, which is the required
  $z$ surface matrix.
  
  This can be acheived with the `data.matrix` function:
  
  
```{r}
z_surface_matrix <- data.matrix(benthic.pred[1])
```

##Base Package
Using base R packages, a contour plot can be created with the `contour`
function:
```{r}
contour(xgrid, ygrid, z_surface_matrix,
          xlab="-Longitude (degrees West)", ylab="Latitude (degrees North)",
          main = "contour plot of Krige Prediction",
          labcex = 0.8)
  mtext("Krige using a 4th degree polynomial model", cex = 0.5)
```


#Part 3: Surface Plots
A surface plot allows a three dimensional view of the model

##Base
In base package, a surface plot can be made:
```{r}
persp(xgrid, ygrid, z_surface_matrix,
        xlim = c(-77.3, -75.9),   #the values domain needs to be adjusted above
        ylim = c(38.1, 39.5),          #Make sure to set an appropriate domain
        # zlim = c(0, 6),
        theta = -45, phi = 30, d = 0.5,
        xlab="-Longitude (degrees West)",
        ylab="Latitude (degrees North)",
        zlab="Benthic Index", ticktype = "detailed")
  title(main=paste("Surface Plot of Benthic Index", "Loess Smoothing", sep="\n"))   
```






















