---
title: "10 Tutorial"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

# Preamble
## Load Packages
```{r}
load.pac <- function() {
  
  if(require("pacman")){
    library(pacman)
  }else{
    install.packages("pacman")
    library(pacman)
  }
  
  pacman::p_load(xts, sp, gstat, ggplot2, rmarkdown, reshape2, ggmap, wesanderson,
                 parallel, dplyr, plotly, tidyverse, reticulate, UsingR, Rmpfr, latex2exp,
                 mise, GGally, usmap)
  
#  devtools::install_github("tidyverse/tidyverse")
}

load.pac()
mise()

```
## Load Data

```{r}
(adult  <- read.csv(file = "./10_data/AdultDataSet/adult.data")) %>%  head()
(aus    <- read.table(file = "./10_data/Australian/australian.dat",  header = TRUE)) %>%  head()
(mob   <- read.csv(file = "./10_data/Datamobile Data Set/imports-85.data")) %>% head()
(house <- read.table(file = "./10_data/Housing/housing.data", header = TRUE) ) %>% head()
```

# Question 1 and 2
## Adult Data
First investigate the Features

```{r}
names(adult)
```


Now plot the data relating to Income Threshold: 
```{r}
ggplot(adult, aes(x = Hours.Per.Week, y = Income.Threshold, fill = Sex)) +
  geom_boxplot() + 
  labs(x = "Hours per Week", y = "Income Threshold", title = "Income Threshold Given Hours per Week Across Genders")
```

this demonstrates that men tend to work longer hours and women need to work longer hours in order to earn above $50K.

```{r}
ggplot(adult, aes(y = Income.Threshold, fill = Sex, x = Wage)) +
  geom_boxplot()
```
This demonstrates that wage has no bearing on Income Threshold, which is likely not true, there could be an issue with the data.

```{r}
ggplot(adult, aes(x = Occupation, fill = Occupation, y = Wage)) +
  geom_col(position = 'dodge') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  guides(fill = FALSE) +
  labs(title = "Occupation and Salary")
```

```{r}
ggplot(adult, aes(x = Education, fill = Education, y = Wage)) +
  geom_col(position = 'dodge') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  guides(fill = FALSE) + 
  labs(title = "Education and Salary")
```

This plot suggests that the less college you complete, the more money you will earn, again, there may be something awry with this data.

## Aus Data
This data set has non descriptive features, for visual analytics this might be better because it allows for analysis without prejudice.

The best thing to do is to use a correlation plot in order to determine which features are the most significant:

```{r}
library(corrplot)
head(aus)
cor(aus) %>% corrplot(method = 'ellipse', type = 'upper')
```

from this we can tell that `X0.3` and `X0` are highly correlated, as are `X0.1` and `X0.2`, there also appers to be weak negative correlation between `x100` and `X11.46`, these can be represented by a scatter plot to better understand the behaviour, for example:

```{r}
ggplot(aus, aes(x = X100, y = X11.46, col = factor(X1))) +
  geom_point() +
  guides(col = guide_legend(TeX("X_1"))) +
  labs(x = TeX("X_{100}"), y = TeX("X_{11.46}"))

```



## DataMobile
The data can be inspected using `str`:

```{r}
str(mob)
```
This reveals the data are car metrics, we can then visualise different aspects of the data that might be interesting:

### Average Price per Manufacturer
```{r}
mob$price <- as.numeric(mob$price)


ggplot(mob, aes(y = price, x = make, fill = make)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  guides(fill = FALSE)
  

```
This provides that Toyota Mercedes and BMW produce some of the most expensive vehicles.

### Vehicle Performance and Fuel Consumption


```{r}

mob$horsepower <- as.numeric(mob$horsepower)

ggplot(mob, aes(x = city.mpg, y = horsepower/curb.weight, col = price)) +
  geom_point() +
  theme_bw() +
  labs(x = TeX("Mileage (\\frac{mi}{gl})"), y = TeX("Power to Weight (\\frac{hp}{lbs})"),
       title = "Vehicle Performance and Fuel Consumption")
```

This indicates that high performance (High Power to Weight) vehicles tend to consume more fuel, this fuel change in fuel consumption appears to scale linearly.

The price appears not to change significantly accross fuel consumption or power to weight.

## House Data
The housing data can be inspected using the `str` function and the information relating to the data set can be found [here](https://stat.ethz.ch/R-manual/R-devel/library/MASS/html/Boston.html)

```{r}
str(house)
```

In order to get an understanding of the relationship between the data a scatter plot matrix can be used:

```{r}
library(GGally)
cor(house) %>% corrplot(method = "ellipse", type = 'upper')

```

This suggests that the median value of the house is signifanctly impacted by the number of rooms `RM` the  lower status of the population  `LSTAT`  and the contentration of nitrous oxides in the air `NOX`, this could be visualised by a scatterplot:

```{r}
ggplot(house, aes(x = NOX, y = MEDV, shape = factor(round(RM)), col = LSTAT))  +
  geom_point() +
  theme_bw() +
  labs(x = TeX("Nitrogen Oxides Concentration (10^{-1} ppm)"),
       y = "Median House Value",
       title = " Housing Prices")
```


# Question 3

```{r}
library("readxl")
(store <- read_excel("./10_data/10_data_Superstore.xls")) %>% str()
names(store)
```


### Sales and Quantity
A plot  of Sales vs Quantity can be produced:

```{r}
store$Quantity <- factor(store$Quantity, levels = 1:max(store$Quantity), ordered = TRUE)
p <- ggplot(data = store, aes(x = factor(Quantity), y = Profit, col = Quantity))
 
  
  
  p + geom_boxplot() +
  guides(col = FALSE) +
  labs(x = "Quantity Units", title = "Profit Given Quantity sold")
```

This suggests that 4/5 units of quantity produce the most variation in profit, for this reason it can be anticipated that higher quantities may be preferred because the profits will be more consistent.

## 2 Post Code and Profits
```{r}
names(store)
ggplot(store, aes(x = "", y = Profit, fill = Region)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar(theta = "y", start = 0)
```
This shows that the South Region earns the most profits.

> Yeah, this is actually wrong, I'm not sure why TBH, I'm going to have to investigate Pie Charts in `ggplot2`, I didn't realise it would be so poorly implemented.

## 3 "Region and Ship Mode"
```{r}
ggplot(store, aes(x = Region, y = Sales, fill = `Ship Mode`)) +
  geom_col(position = 'dodge')
```
This indicates that most sales are made in the South and Central Regions and that the shipping method that produces the most profits is *Standard Class*.

## 4 Segment and State

```{r}
ggplot(data = store, aes(x = State, y = Profit, fill = Segment)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

This provides that New York and and Arkansas were the states that had the biggest impact on profits, generally consumer purchases generate the most profits.


## 5 Quantity and Discount

```{r}
ggplot(data = store, aes(x = Discount, y = Quantity)) +
  geom_col(col = "royalblue")
```

This suggests that the largest discount is given for 2/3 items where as purchasing many items does not often involve a discount.

## Trying to Map States
This is as far as I got, the issue is that I'd need to convert the data to have lat/long or standard names which would take a bit of research, whereas *Tablaeu* can interpret US names in a way that hasn't been implemented in `ggplot2`.

```{r}
library(usmap)
library(ggplot2)

plot_usmap(data = statepop, values = "pop_2015", color = "red") + 
  scale_fill_continuous(name = "Population (2015)", label = scales::comma) + 
  theme(legend.position = "right") 
```

# Question 4
Different colour pallettes could have been used in these visualisations, for example:

## R Colour Brewer Discrete

```{r}
ggplot(store, aes(x = Region, y = Sales, fill = `Ship Mode`)) +
  geom_col(position = 'dodge') +
  scale_fill_brewer(palette = 'Dark2')


```

## Custom Discrete Pallete

```{r}
pt <- c("#AE60B8", "#ED2B78", "#B8E3B8", "#FA9F7B")
ggplot(store, aes(x = Region, y = Sales, fill = `Ship Mode`)) +
  geom_col(position = 'dodge') +
  scale_fill_manual(values = pt)

```

## Continuous Pallete

```{r}
ggplot(house, aes(x = NOX, y = MEDV, shape = factor(round(RM)), col = LSTAT))  +
  geom_point() +
  theme_bw() +
  labs(x = TeX("Nitrogen Oxides Concentration (10^{-1} ppm)"),
       y = "Median House Value",
       title = " Housing Prices") +
  scale_colour_gradient(low = "green", high = "darkblue")
```




















































