---
title: "Exercise 12.3"
output: 
  html_notebook: 
    toc: yes
---

#Abstract
The goal here is to create a prediction model for 
the lead level over the Meuse River region

This prediction model will be graphically portrayed with a:

1. Heatmap
    i) with base plot
    ii) wigh `ggplot2` over an actual map
2. Contour plot
    i) with base plot
    ii) With `ggplot2` over an actual map
3. 3d Surface plot
    i) with base plot
    ii) with a dynamic `plotly` surface plot

##Log-Transformation
The Log transformation is used to make the data
more normally distributed, it is required because
it creates predicted values that more closely resemble the
the observed values.
##Lead-Levels
It can be seen that there are pockets of high lead levels on the
shore of the river at three points.

#Preamble
##Import Library

The following code will automatically install packages that are 
not already installed and load them into the library:

```{r}
if(require('pacman')){
  library('pacman')
}else{
  install.packages('pacman')
  library('pacman')
}

pacman::p_load(sp, EnvStats, gstat, ggplot2, rmarkdown, reshape2, ggmap,
               RColorBrewer, parallel, dplyr, plotly, proj4, rgdal)
```


##Create Assignments

Note that  meuse data set is included in the `sp` package and is not a 
data frame but an `sp` class, which is very similar to a data frame

```{r}
#Inspect the Data Set
data(meuse) 
head(meuse)

#Assignments
lon <- meuse$x
lat <- meuse$y
cd <- meuse$cadmium
cu <- meuse$copper
pb <- meuse$lead
zn <- meuse$zinc

```


#Lattitude and Longitude
* Lattitude is a line that runs east to west that measures how far 
north of the equatour a location is
      + Lattitude is assigned to the y-axis
* Longitude is a line that runs north to south that measures east/west
     + Longitude is assigned to the x-axis



#Assign the Co-ordinates
The *meuse* Data set is an `sp` class objects and requires for its co-ordinates to be set:

```{r}
rm(meuse); data(meuse)
coordinates(meuse) = ~x + y
```

The co-ordinate reference system is in meters north and meters east,
this is going to be incompatible with the `ggmap` function, so
it will have to be transformed first.

Because we are only plotting with base packages, pursuant
to the lecture notes, we will leave the co-ordinate system in
eastings and northings.





#Preliminary Step: Create a krige prediction Model
In order to create a krige prediction model two things are required:

* Fitted Variogram Model (e.g. an exponential model)
* A domain values over which to predict values

##Create a variogram Model


###Create a Variogram 
Using an ordinary kriging method without any surface model provided:

```{r}

lead.vg <- variogram(object = lead~1, data = meuse)
plot(lead.vg)
head(lead.vg)
```


###Fit an exponential model to that variogram

```{r}
#This is what would be used for the northing/easting CRS
lead.vg.fit <- fit.variogram(lead.vg, model = vgm(1, "Sph", 900, 1)) 


```

Thus the variogram model needed for the krige prediction has been ascertained

##Create a Grid of Domain values to use as predictor values
### Create a sequence of incrementally increasing for the X and Y axis
This will essentially act as the resolution of the overlay,
Most monitors are about 140 ppi, hence assuming the plot will be a
9 inch square, about 800 x 800 pixels should look right, but 
200 x 200 will be the used due to performance.

```{r}
xgrid       <-  seq(min(lon), max(lon), length.out = 200)
ygrid       <-  seq(min(lat), max(lat), length.out = 200)
```


###Combine into a domain base
This domain base however has to be a spatial object, this can be
acheived by defining co-ordinates in the data frame
```{r}
xy.surface  <- expand.grid(lon = xgrid, lat = ygrid)
coordinates(xy.surface) = ~lon + lat
proj4string(xy.surface) <- proj4string(meuse)
head(xy.surface, 3)
```

##Create the Krige prediction object
Krige predictions can be more accurate if the data is normally 
distributed.

Hence the case study used a log transformation on the *Zn* value, this is because
the *Zn* value was non-normally distributed and the log transformation
somewhat fixed that.

```{r}
par(mfcol = (c(1,2)))
hist(zn, main = "Zinc Value")
hist(log(zn), main = "Log-Transformation of Zinc Value")
```

An inspection of the *Pb* values shows that the log-transformation
may be somewhat more normal, hence it will be used in the analysis.

```{r}
par(mfcol = (c(1,2)))
hist(pb, main = "Lead Value")
hist(log(pb), main = "Log-Transformation of lead Value")
```



```{r}

lead.pred <- krige(formula = lead ~1,
                           locations =  meuse,
                           newdata = xy.surface,
                           model= lead.vg.fit.exp)


lead.log.pred = krige(log(lead)~1, meuse, xy.surface, model = lead.vg.fit)
lead.pred = krige(lead~1, meuse, xy.surface, model = lead.vg.fit)

par(mfcol = c(2,2))
hist(pb, main = "Lead Values",
     xlab = "Lead (ppm)")
hist(lead.pred$var1.pred, main = "Prediction using Lead Values",
     xlab = "Lead (ppm)")
hist(pb, main = "Lead Values",
     xlab = "Lead (ppm)")
hist(exp(lead.log.pred$var1.pred),
     main = "Prediction using Log Transformation",
     xlab = "Lead (ppm)")

```

It can be seen that the prediction using the log transformation
has a shape that (ever so slightly) more closely resembles the distribution of lead values,
hence it is used:

```{r}

lead.log.pred$var1.pred <- exp(lead.log.pred$var1.pred)

head(lead.log.pred)

```

Now there is a prediction model, if that prediction model is
extrapolated over our domain (`xy.surface`) the various plots can be
created

#Part 1: Create a heatmap

In order to create a heatmap, it is first necessary to take
the `sp` prediction object and make it into pixel data,
otherwise the heatmap won't look quite right:

```{r}
class(lead.pred)
gridded(lead.pred) = TRUE
class(lead.pred)
```


```{r}
spplot(lead.pred["var1.pred"],main="Kriging Predictions of lead levels (ppm) in Meuse River")
```


#Part 2: Contour Map
In order to view the surface plot, contours can also be used.

 Recall from Week 11/Exercise 10, that contour and surface plots
  work a little differently to ordinary plots.
  
  A domain grid of equally spaced $x$ and $y$ values is required and
  a matrix of $z$ values that can be overlayed onto that domain grid
  in order to provide the surface values required 
  to create the surface plot.



  The method from Lecture 10/Wk. 11 involved predicting over
  the square domain of `xy.surface`, this won't work here
  because instead of the `predict` function, the `krige` function
  was used, hence it is necessary to convert the output from
  the `krige` function, into the output that would have been 
  given by the predict function, which is the required
  $z$ surface matrix.
  
  This can be acheived with the `data.matrix` function:
  
  
```{r}
z_surface_matrix <- data.matrix(lead.pred[1])
```

##Base Package
Using base R packages, a contour plot can be created with the `contour`
function:
```{r}
contour(xgrid, ygrid, z_surface_matrix,
          xlab="-Longitude (degrees West)", ylab="Latitude (degrees North)",
          main = "contour plot of Krige Prediction",
          labcex = 0.8)
  mtext("Krige using a 4th degree polynomial model", cex = 0.5)
```

#Part 3: Surface Plots
A surface plot allows a three dimensional view of the model

##Base
In base package, a surface plot can be made:
```{r}
persp(xgrid, ygrid, z_surface_matrix,
        theta = -45, phi = 30, d = 0.5,
        xlab="-Longitude (degrees West)",
        ylab="Latitude (degrees North)",
        zlab="Lead Levels (ppm)", ticktype = "detailed")
  title(main=paste("Surface Plot of Lead Levels (ppm)", "Loess Smoothing", sep="\n"))   
```
##Plotly
Although plotly is not a base package, it is easier to use, so consider using
it instead