---
title: "Exercise 12.1"
output: 
  word_document: 
    toc: yes
  html_notebook: 
    toc: yes
---
The goal here is to create a prediction model for 
the value of benthic over the Chesepeake bay region.

This prediction model will be graphically portrayed with a:

1. Heatmap
    i) with base plot
    ii) wigh `ggplot2` over an actual map
2. Contour plot
    i) with base plot
    ii) With `ggplot2` over an actual map
3. 3d Surface plot
    i) with base plot
    ii) with a dynamic `plotly` surface plot


#Preamble
##Import Library

The following code will automatically install packages that are 
not already installed and load them into the library:

```{r}
if(require('pacman')){
  library('pacman')
}else{
  install.packages('pacman')
  library('pacman')
}

pacman::p_load(sp, EnvStats, gstat, ggplot2, rmarkdown, reshape2, ggmap, RColorBrewer, parallel, dplyr, plotly)
```

##Create Assignments

Note that  Benthic.df is included in the `EnvStats` package and is not a 
data frame but an `sp` class, which is very similar to a data frame

```{r}
#Assignments
benthic <- Benthic.df
lat   <- benthic$Latitude
lon   <- benthic$Longitude
index <- benthic$Index
```

#Inspect the Data Set

```{r}

head(benthic, 3)
```

#Lattitude and Longitude
* Lattitude is a line that runs east to west that measures how far 
north of the equatour a location is
      + Lattitude is assigned to the y-axis
* Longitude is a line that runs north to south that measures east/west
     + Longitude is assigned to the x-axis



#Assign the Co-ordinates
The *Benthic* Data set is an `sp` class objects and requires for its co-ordinates to be set:

```{r}
coordinates(benthic) = ~Longitude + Latitude
```
#Preliminary Step: Create a krige prediction Model
In order to create a krige prediction model two things are required:

* Fitted Variogram Model (e.g. an exponential model)
* A domain values over which to predict values

##Create a variogram Model




###Create a Variogram 
If a 4th degree polynomial model was used to model the 
benthic values over space, a variogram can be made from
that model thusly:

(The quadratic model is a required parameter of the kriging process,
in exercise 12.2 a quadratic model will be used,
the spacific mathematics of this though is somewhat beyond me, however
the kriging method uses the residuals from that surface model
in order to make the predictions.)

```{r}
x <- lon
y <- lat

benthic.vg <- variogram(object = index ~ 
                          x + y + 
                          x^2 + x*y + y^2 + 
                          x^3 + x^2*y + x*y^2 + y^3 +
                          x^4 + x^3*y + x^2*y^2 + x*y^3 + y^4,
                        data = benthic)

head(benthic.vg)
```


###Fit an exponential model to that variogram

```{r}
benthic.vg.fit.exp <- fit.variogram(benthic.vg, model=vgm(1,"Exp", 0.5,1))
```

Thus the variogram model needed for the krige prediction has been ascertained

##Create a Grid of Domain values to use as predictor values
### Create a sequence of incrementally increasing for the X and Y axis
This will essentially act as the resolution of the overlay,
Most monitors are about 140 ppi, hence assuming the plot will be a
9 inch square, about 800 x 800 pixels should look right, but 
200 x 200 will be the used due to performance.
```{r}
xgrid       <-  seq(min(lon), max(lon), length.out = 40)
ygrid       <-  seq(min(lat), max(lat), length.out = 40)
```


###Combine into a domain base
This domain base however has to be a spatial object, this can be
acheived by defining co-ordinates in the data frame
```{r}
xy.surface  <- expand.grid(lon = xgrid, lat = ygrid)
coordinates(xy.surface) = ~lon + lat
head(xy.surface, 3)
```

##Create the Krige prediction object
```{r}

benthic.pred <- krige(formula = Index ~1,
                           locations =  benthic,
                           newdata = xy.surface,
                           model= benthic.vg.fit.exp)
head(benthic.pred)
```

Now there is a prediction model, if that prediction model is
extrapolated over our domain (`xy.surface`) the various plots can be
created

#Part 1: Create a heatmap

In order to create a heatmap, it is first necessary to take
the `sp` prediction object and make it into pixel data,
otherwise the heatmap won't look quite right:

```{r}
class(benthic.pred)
gridded(benthic.pred) = TRUE
class(benthic.pred)
```


```{r}
spplot(benthic.pred["var1.pred"],main="Predictions of Benthic Index - ordinary kriging")
```


##Create it in ggplot2
First it is necessary to create a data frame from the `sp` object:

```{r}
benthic.pred.df      <- cbind(benthic.pred@coords, benthic.pred@data)
benthic.pred.df.tidy <- melt(benthic.pred.df)
head(benthic.pred.df)

ggplot() +
  geom_tile(data = benthic.pred.df, aes(lon, lat, fill = var1.pred)) +
  labs(fill = "Benthic Index") # 
```



This plot would be easier to read if it had discrete intervals
rather than a continuous scale:

```{r}
benthic.pred.df$bin <- cut(benthic.pred.df$var1.pred,
                            breaks = as.vector(seq(from = 0, to = 5, length.out = 20)))

ggplot() +
  geom_tile(data = benthic.pred.df, aes(lon, lat, fill = bin)) +
  labs(fill = "Benthic Index", 
       title = "Heatmap of Benthic Index Values",
       x = "Longitude",
       y = "Lattitude")
  
```

###Map overlay

Now by adjusting the colour scale and opacity, this can be used as an overlay
for the map from exercise 10.1:

```{r}
benthic.pred.df$bin <- cut(benthic.pred.df$var1.pred,
                            breaks = as.vector(seq(from = 0, to = 5, length.out = 9)))

#Get a map
  bbox <- make_bbox(benthic.pred.df$lon, benthic.pred.df$lat, f = 0.01)
  map <- get_map(location = bbox, maptype = 'toner')
  #Create the colours
  cols <- brewer.pal(n = length(levels(benthic.pred.df$bin)), name = "Oranges") 
  #Create the Plot
  ggmap(ggmap = map) +
  geom_tile(data = benthic.pred.df, aes(lon, lat, fill = bin, alpha = var1.pred))+
  scale_fill_manual(values = cols) +
  labs(fill = "Benthic Index",
       x = "Longitude",
       y = "Lattitude", 
       title = "Heatmap of Benthic Index") + 
  guides(alpha=FALSE) 
    
```


#Part 2: Contour Map
In order to view the surface plot, contours can also be used.

 Recall from Week 11/Exercise 10, that contour and surface plots
  work a little differently to ordinary plots.
  
  A domain grid of equally spaced $x$ and $y$ values is required and
  a matrix of $z$ values that can be overlayed onto that domain grid
  in order to provide the surface values required 
  to create the surface plot.



  The method from Lecture 10/Wk. 11 involved predicting over
  the square domain of `xy.surface`, this won't work here
  because instead of the `predict` function, the `krige` function
  was used, hence it is necessary to convert the output from
  the `krige` function, into the output that would have been 
  given by the predict function, which is the required
  $z$ surface matrix.
  
  This can be acheived with the `data.matrix` function:
  
  
```{r}
z_surface_matrix <- data.matrix(benthic.pred[1])
```

##Base Package
Using base R packages, a contour plot can be created with the `contour`
function:
```{r}
contour(xgrid, ygrid, z_surface_matrix,
          xlab="-Longitude (degrees West)", ylab="Latitude (degrees North)",
          main = "contour plot of Krige Prediction",
          labcex = 0.8)
  mtext("Krige using a 4th degree polynomial model", cex = 0.5)
```

##Using ggplot2
In order to create a contour plot in ggplot 2,
it will be necessary to use 'tidy' data, which is 
a table of data such that each column corresponds
to a variable:
```{r}
rownames(z_surface_matrix) <- xgrid
  colnames(z_surface_matrix) <- ygrid
  benthic.pred.melt <- melt(z_surface_matrix,
                            varnames = c("Longitude", "Latitude"),
                            value.name = "Benthic_Index")
  head(benthic.pred.melt)
```
  Observe that this is another way of getting the
  data frame that was required in part 1 (`benthic.pred.df`),
  it is included here as another method and for the sake of completion,
  also it matches the method used in week 10/lecture 11.


Now using ggplot2 a contour plot, can be created:

```{r}
ggplot(data = benthic.pred.melt, aes(x = Longitude,
                                       y = Latitude,
                                       z = Benthic_Index,
                                       colour = ..level..),
         size = 0.5) +
    labs(x = "Longitude", y = 'Latitude',
         title = 'Benthic Index',
         subtitle = "Chesapeake Bay, Maryland",
         caption = "Higher index value is better",
         col = 'Benthic Index', size = 'Benthic Index') + 
    geom_contour(lwd = 1) +
  theme_classic()
```

###Map Overlay
The advantage to using ggplot2, is the map overlay can be taken advantage of:


```{r}
  map <- get_map(location = bbox, maptype = 'watercolor')
  
  ggmap(ggmap = map) +
    labs(fill = "Benthic Index",
         x = "Longitude",
         y = "Lattitude", 
         title = "Heatmap of Benthic Index") + 
    guides(alpha=FALSE) +
    geom_contour(data = benthic.pred.melt, aes(x = Longitude,
                                               y = Latitude,
                                               z = Benthic_Index),
                 col = 'grey40',
                 binwidth = 0.15,
                 size = 0.70) +
    scale_fill_manual(values = cols) 
```

###Map Overlay with Heatmap
This can be combined with the heatmap layer:

```{r}
  ggmap(ggmap = map) +
    labs(fill = "Benthic Index",
         x = "Longitude",
         y = "Lattitude", 
         title = "Heatmap of Benthic Index") + 
    guides(alpha=FALSE) +
    geom_tile(data = benthic.pred.df, aes(lon, lat, fill = bin, alpha = var1.pred)) +
    geom_contour(data = benthic.pred.melt, aes(x = Longitude,
                                               y = Latitude,
                                               z = Benthic_Index),
                col = 'indianred',
                binwidth = 0.2,
                size = 0.6) +
    scale_fill_manual(values = cols) 
```

#Part 3: Surface Plots
A surface plot allows a three dimensional view of the model

##Base
In base package, a surface plot can be made:
```{r}
persp(xgrid, ygrid, z_surface_matrix,
        xlim = c(-77.3, -75.9),   #the values domain needs to be adjusted above
        ylim = c(38.1, 39.5),          #Make sure to set an appropriate domain
        # zlim = c(0, 6),
        theta = -45, phi = 30, d = 0.5,
        xlab="-Longitude (degrees West)",
        ylab="Latitude (degrees North)",
        zlab="Benthic Index", ticktype = "detailed")
  title(main=paste("Surface Plot of Benthic Index", "Loess Smoothing", sep="\n"))   
```

##Plotly
Plotly has the advantage of being fully interactive and not as difficult to 
constrain within a plot/box area:
```{r}
p <- plot_ly(x = ygrid, y = xgrid, z = z_surface_matrix) %>%
    add_surface() %>%
    layout(
      title = "Benthic Index in  Chesapeake Bay",
      scene = list(
        xaxis = list(title = "Longitude",
                     range = c(38.2, 38.9)),  #You shouldn't need to edit the
        yaxis = list(title = "Latitude",      #The domain here, do it above
                     range = c(-77, -75.9)),
        zaxis = list(title = "Benthic Index")
      ))
                                                                  
  #p
    #This output is HTML, it cannot be rendered inside a word document
```



The surface plot really reinforces the fact that the bay has
the lowest (i.e. best) benthic index value.


























