---
title: "2020 Project"
subtitle: "Thinking About Data"
author: "Ryan G - 17805315"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---


# Declaration
* I hold a copy of this assignment that we can produce if the original is lost or damaged.
* I hereby certify that no part of this assignment/product has been copied from any other student’s work
or from any other source except where due acknowledgement is made in the assignment.
* No part of this assignment/product has been written/produced for us by another person except where
such collaboration has been authorised by the subject lecturer/tutor concerned.
* I am aware that this work may be reproduced and submitted to plagiarism detection software programs
for the purpose of detecting possible plagiarism (which may retain a copy on its database for future
plagiarism checking).
* I hereby certify that we have read and understand what the School of Computing and Mathematics
defines as minor and substantial breaches of misconduct as outlined in the learning guide for this unit.

# Problem
>  The health minister of Australia has been concerned about the increase in new cases of people infected with
> the Tiara virus. They have hired you as a consultant to examine their data on the new cases of Tiara virus
> and the flu. The data is provided to you in a CSV file called project2020A.csv containing observations of
> the daily new cases of both the Tiara and flu viruses for each state in Australia over a 100 day period. The
> variables are:
> 
> * `city`: The city in which the new cases of the virus is counted.
> * `newTiara`: The number of new cases of the Tiara virus for the day.
> * `newFlu`: The number of new cases of the flu virus for the day.
> * `date`: the day number starting from 1 and ending at 100.
> 
> The health minister wants statistics measured from the data so that they can be reported to the public. The
> five required pieces of analysis are presented below.
# Preamble
Before beginning it is necessary to set the working directory, load any necessary packages and load the data set.


```{r}
## Preamble

## Install Pacman
load.pac <- function() {
  
  if(require("pacman")){
    library(pacman)
  }else{
    install.packages("pacman")
    library(pacman)
  }
  
    pacman::p_load(xts, sp, gstat, ggplot2, rmarkdown, reshape2, ggmap,
                 parallel, dplyr, plotly, tidyverse, reticulate, UsingR, Rmpfr,
                 swirl, corrplot, gridExtra, mise, latex2exp, tree, rpart, lattice,
                 coin, primes, epitools, maps, clipr, ggmap, RColorBrewer)
  
  
   mise()
   select <- dplyr::select
}

load.pac()
setwd(dir = "/home/ryan/Notes/DataSci/ThinkingAboutData/")
load(file = "~/Notes/DataSci/ThinkingAboutData/TAD.rdata")
load(file = "./TAD.rdata")
print("Success")
```

```{r}
data <- read.csv("../0datasets/project2020A.csv")
```

# question 1
## Plot
## Observations from Plot
## Analysis and Results
## Conclusion
# Question 2
## Plot
## Observations from Plot
## Analysis and Results
## Conclusion
# Question 3
## Plot
## Observations from Plot
## Analysis and Results
## Conclusion

# Question 4
## Plot
## Observations from Plot
## Analysis and Results
## Conclusion
# Question 5
## Plot
## Observations from Plot
## Analysis and Results
## Conclusion
