---
title: "Exercise 11.1"
output: 
  html_document: 
    toc: yes
  html_notebook: 
    highlight: tango
    theme: journal
  pdf_document: 
    toc: yes
  word_document: 
    toc: yes
---

#Preamble
##Import Library

The following code will automatically install packages that are 
not already installed and load them into the library:

```{r}
if(require('pacman')){
  library('pacman')
}else{
  install.packages('pacman')
  library('pacman')
}

pacman::p_load(sp, EnvStats, gstat, ggplot2, rmarkdown)
```

##Create Assignments

Note that  Benthic.df is included in the *EnvStats* package and is not a 
data frame but an *sp* class, which is very similar to a data frame

```{r}
#Assignments
benthic <- Benthic.df
lat <- benthic$Latitude
lon <- benthic$Longitude
```

#Inspect the Data Set

```{r}
head(benthic, 3)
```

#Asign Co-ordinates
The benthic data set is not a mere data frame, it is an `sp` object,
which is very similar to a data frame.
It is necessary to set the co-ordinates of this object though:
```{r}
coordinates(benthic) = ~Longitude + Latitude
```
#Create a variogram
A variogram is a description of the structure of spacial data,
first this needs to be created before anything can 
be plotted or modelled:

```{r}
benthic.vg <- variogram(Index ~ 1, data = benthic)
```


Now we can inspect the variogram:
```{r}
head(benthic.vg)
```
By observing this, we can conclude, that a variogram will have an
$x$-axis of `dist` and a $y$-axis of `gamma`.


#Plot the Variogram
##Create a base plot of that variogram

```{r}
plot(benthic.vg, main="Figure 11.1  Empirical variogram for Benthic Index",
     col = 'violetred4', cex = 2)
```

##Create an attractive plot in ggplot
The `ggplot2` package makes it reasonable easy to
add a non-parametric model using the $loess$ method:

```{r}
ggplot(data = benthic.vg, aes(x = dist, y = gamma, col = gamma, size = np)) +
  geom_point() +
  scale_color_continuous(low = 'royalblue', high = 'indianred') +
  labs(size = "Number of Point Pairs", col = "Sample Variogram Estimate",
       y = "Sample Variogram estimate", x = "Distance",
       title = 'Benthic Variogram Plot') +
  theme_bw() +
  geom_smooth(se = TRUE, fill = 'purple', alpha = 0,
              size = 1, span = 2, col = 'indianred2')


```


#Create Variogram Models
Variogram models are objects of the class `variogramModel` and 
cannot be drawn with `ggplot2`, hence only base plots can
be used to draw a parametric model over the variogram plot.

## Exponential Model
An exponential model is of the form:

$C(h) = \sigma^2 \times e^{\frac{-h}{r}}$

and can be modelled:
```{r}
benthic.vg.fit.exp <- fit.variogram(benthic.vg,
                                    model=vgm(1,"Exp", 0.5,1))
plot(benthic.vg, benthic.vg.fit.exp,
     main="Exponential variogram for Benthic Index")
```

#Gaussian Model:
A Gaussian model is of the form:

$C(h) = \sigma^2 \times e^{({\frac{-h}{r}})^2}$

and can be modelled:
```{r}

benthic.vg.fit.gau <- fit.variogram(benthic.vg,
                                    model=vgm(1,"Gau", 0.5,1))
plot(benthic.vg, benthic.vg.fit.gau,
     main="Gaussian variogram for Benthic Index")
```

#Spherical Model:
A sperical model is of the form:

$C(h) = \sigma^2 \times (1 - \frac{3h}{2r} + \frac{h^2}{2r^3})$


```{r}

benthic.vg.fit.sph <- fit.variogram(benthic.vg,
                                    model=vgm(1,"Sph", 0.5,1))
plot(benthic.vg, benthic.vg.fit.sph,
     main="Spherical variogram for Benthic Index")
```

#Directional Variogram

Because a directional variogram is still of the class `variogramModel`, 
it is not possible (to the best of my knowledge) to use `ggplot2`
in order to plot it.

The `alpha` option allows a vector of directional degrees to be specified, in this case, 0, 45, 90 and 135.

```{r}
benthic.dvg <- variogram(Index ~ 1,
                         data=benthic, alpha=c(0, 45,90,135))
benthic.dvg.fit <- vgm(1,"Sph", 0.5,1,anis=c(30,0.4))
plot(benthic.dvg, benthic.dvg.fit,
     main="Figure 11.5  Directional variograms for Benthic Index")
```



















