---
title: "04 Binomial Distribution and Matrices"
tags: [Notebooks/Analytics Programming]
# output: 
#   html_document: 
#     keep_md: yes
output: 
  html_notebook: 
    toc: yes
    toc_depth: 5
  pdf_document: 
    latex_engine: pdflatex # xelatex is really slow
    pandoc_args: "--pdf-engine-opt=-shell-escape"
    toc: yes
    toc_depth: 5
header-includes:
  - \usepackage{\string~/Dropbox/profiles/Templates/LaTeX/ScreenStyle}
---


# Practical 04


## Preamble

```{r}

## (01) Clean up the Iris Data

# Preamble
## Install Pacman
load.pac <- function() {
  
  if(require("pacman")){
    library(pacman)
  }else{
    install.packages("pacman")
    library(pacman)
  }
  
  pacman::p_load(xts, sp, gstat, ggplot2, rmarkdown, reshape2, ggmap,
                 parallel, dplyr, plotly, tidyverse, reticulate, UsingR, Rmpfr,
                 swirl, corrplot, gridExtra, mise, latex2exp, tree, rpart)
  
}

load.pac()
 mise()
```

For the project it will be necessary to:

1. Generate Random Binomial Distributions
2. View elements of a matrix
3. Use:
   * Loops, and
   * Conditional Statements
4. Visualise Results

## (01) Binomial Distribution and Viewing Elements in a Matrix

### Create Binomial Values

```{r}
n <- 10 # Matrix Size
binom_values <- rbinom(n^2, size = 20, prob = 0.45)
```

### Create a matrix of Binomial Values

```{r}
(starting_point <- matrix(binom_values, nrow = 10))
```

#### Looping through a matrix

```{r}


for (r in 1:nrow(starting_point)) {
  for (c in 1:ncol(starting_point)) {
      cat(starting_point[r,c])
  }
 cat("\n")
}


x <- 33


print(paste("x is ", x))
cat(paste("x is ", x))

library(tidyverse)
paste("x is ", x) %>% print()
```



## (02) Visualise Data Based on a Condition

### Create the Conditional Matrix

In order to visualise a matrix based on a condition it is necessary to create a truth matrix, there are two ways to do this, the ordinary way:

```{r}
zero <- starting_point>10
```

and with nested `for` loops, which may be necessary for some certain tasks:

```{r}
for (r in 1:nrow(starting_point)) {
  for (c in 1:nrow(starting_point)) {
    if (starting_point[r, c] < 10) {
      zero[r, c] <- 1
    }
  }
}
zero
```

### Plot the Conditional Matrix

#### Base Plot

The base plot involves using `par` to set graphical parameters to to initialise a blank canvas, rotating the matrix 90 degrees clockwise and then passing that rotated matrix to the `image()` function:

```{r}
# Create a Plotting Region
par(pty = "s", mai = c(0.1, 0.1, 0.4, 0.1))

# Rotate matrix 90-degrees clockwise.
rotate <- function(x) {
    t(apply(x, 2, rev))
}

# create the image
image(rotate(zero), col = c(3, 6), axes = FALSE, frame.plot = TRUE, main = "Values > 10")


image(rotate(zero), col = c("red", "black"), axes = FALSE, frame.plot = TRUE, main = "Values > 10")
```


##### Rotation
It is necessary to rotate the matrix 90 degrees clockwise before passing it to the `image()` function in order for it to line up correctly, this is achieved by reversing the order of the columns (rotating around a horizontal line) and then rotating around the line y=x or `#1 = #2`.


```{r}
vals <- c()

for (r in 1:5) {
    for (c in 1:5) {
        vals <- c(vals, paste0(c, ",", r))
    }
}

(dummy_mat <- matrix(vals, nrow = sqrt(length(vals))))
rotate(dummy_mat)
```

The image function will allocate values along the cartesian plane:

* from (0, 0) along the $x$ - axis and then
* from (0, 1) along the $x$ -axis and then 
* from (0, 2) along the $x$ -axis and so on
* ...

but the values will be read as vertical columns (say the matrix has 5 rows as above):

* from [1,1] to [5, 1]
* from [2,1] to [5, 2]
* from [3,1] to [5, 3]
* ...

This has the effect of the matrix being plotted at a rotation of 90 degrees ACW and hence the input data must be rotated 90 degrees CW.


#### Using `ggplot2` 

Generally to make a heatmat in ggplot2, use the following syntax:

```{r}
set.seed(8398)
x       <- LETTERS[1:10]
y       <- paste("var", 1:10)
data    <- expand.grid("x" = x, "y" = y)
data$z  <- factor(sample(1:2, 100, replace = TRUE))
data %>% head()

ggplot(data, aes(x = x, y = y, fill =z)) +
  geom_tile()
```


The `expand.grid` function basically just treats `x` and `y` as the axis of a data Frame that is then transformed into a *longer* or *tidy* format.

The benefit to using `expand.grid` is that `x` and `y` will always be arranged as they would be on a cartesian plane and then the data is transformed into a tidy format.

You'll notice, if you inspect the output of `expand.grid` that the values are entered into the data frame from 'bottom-to-top, left-to-right', hence if:

* the rows of a matrix are considered the $y$ -axis, and
* the columns the $x$ -axis 

The values of the matrix would feed into the expand grid row-wise, starting from the bottom moving up, **_R_** converts matrices to vectors column-wise from left to right, so it would be necessary to rotate the matrix [90 degrees clockwise](https://www.geeksforgeeks.org/rotate-a-matrix-by-90-degree-in-clockwise-direction-without-using-any-extra-space/), this is what's done by the `rotate()` function.

When using `ggplot2` it is necessary to make sure that row and column names are ordered factors otherwise they will be plotted incorectly (i.e as floating points or out of order), So first use the `paste()` function to create nice column and row names:

```{r}
# Functions to Name the Rows and Columns
make_colnames <- function(mat) {
                     factor(paste("col", 1:ncol(mat)),
                            levels = paste("col", 1:ncol(mat)),
                            ordered = TRUE)
}
make_rownames <- function(mat) {
                     factor(paste("row", 1:nrow(mat)),
                            levels = paste("row", 1:nrow(mat)),
                            ordered = TRUE)
}

make_rownames_rev <- function(mat) {
                     factor(paste("row", nrow(mat):1),
                            levels = paste("row", nrow(mat):1),
                            ordered = TRUE)
}

make_ynames <- function(mat) {
                     factor(paste("y=", 1:nrow(mat)),
                            levels = paste("y=", 1:nrow(mat)),
                            ordered = TRUE)
}
make_xnames <- function(mat) {
                     factor(paste("x=", 1:ncol(mat)),
                            levels = paste("x=", 1:nrow(mat)),
                            ordered = TRUE)
}



rownames(zero) <- make_rownames(zero)
colnames(zero) <- make_colnames(zero)
```

Now plot the matrix using `expand.grid()`

```{r}
# Create a Cartesian Plane of column names and the rownames backwards
    # This function will then convert that table into tidy format.
data <- expand.grid("x" = make_colnames(zero), "y" = make_rownames_rev(zero))
# Rotate the Matrix 90 degrees clockwise
data$z <- as.vector(rotate(zero))

# Print the Plotted Data
zero
ggplot(data, aes(x = x, y = y, fill =z)) +
  geom_tile()
```

##### As a Function

All together, functions to plot a matrix as a heatmap are:

```{r}


# Rotate matrix 90-degrees clockwise.
rotate <- function(x) {
    t(apply(x, 2, rev))
}

rotate_ACW <- function(x) {
    t(apply(x, 1, rev))
}
  


  # Create Column Names  
make_colnames <- function(matrL) {
                     factor(paste("col", 1:ncol(matrL)),
                            levels = paste("col", 1:ncol(matrL)),
                            ordered = TRUE)
}

 # Create Row Names
make_rownames <- function(matrL) {
                     factor(paste("row", 1:nrow(matrL)),
                            levels = paste("row", 1:nrow(matrL)),
                            ordered = TRUE)
}

# Create Row Names with reversed Order
make_rownames <- function(matrL) {
                     factor(paste("row", nrow(matrL):1),
                            levels = paste("row", nrow(matrL):1),
                            ordered = TRUE)
}


expand.grid_matrix <- function(mat) {
  data <- expand.grid("x" = make_colnames(mat), "y" = make_rownames_rev(mat))
  data$z <- as.vector(rotate(mat))
  
  return(data)
}

(unsquare_test_matrix <- matrix(factor(sample(1:3, 36, replace = TRUE)), nrow = 9))
data <- expand.grid_matrix(unsquare_test_matrix)

ggplot(data, aes(x = x, y = y, fill =z)) +
  geom_tile()
```



##### Use Data Frames (not expand.grid())
Given that the data is already in the form of a matrix, it may be more convenient (or make the plotting process easier to understand) to work with data frames, in order to do this:

  1. Reverse the direction of the columns so that columns are numbered from bottom to top (as they would be on a cartesian plane).
  2. convert the matrix to a dataframe / tibble
     + Make sure that any names are ordered factors, otherwise ggplot will rearrange them into alphabetical order.
  3. Transform the data into `long` format using `pivot_longer`.

So first use the `paste()` function to create nice column and row names:

```{r}
# Functions to Name the Rows and Columns
make_colnames <- function(mat) {
                     factor(paste("col", 1:ncol(mat)),
                            levels = paste("col", 1:ncol(mat)),
                            ordered = TRUE)
}
make_rownames <- function(mat) {
                     factor(paste("row", 1:nrow(mat)),
                            levels = paste("row", 1:nrow(mat)),
                            ordered = TRUE)
}

make_ynames <- function(mat) {
                     factor(paste("y=", 1:nrow(mat)),
                            levels = paste("y=", 1:nrow(mat)),
                            ordered = TRUE)
}
make_xnames <- function(mat) {
                     factor(paste("x=", 1:ncol(mat)),
                            levels = paste("x=", 1:nrow(mat)),
                            ordered = TRUE)
}
```
  
  Now implement steps 1 to 3; reverse the matrix direction, create a data frame and finally make a tidy data frame:

```{r}
set.seed(9832832)
zero <- zero==1

# 0. Name the Rows and Columns
  colnames(zero) <- make_colnames(zero)
  rownames(zero) <- make_rownames(zero)
    ## Row Names are reversed relative to y-axis values

# 1. Reverse the Columns
  zero_rev_col <- apply(zero, 2, rev)

# 2. Create a Data Frame
  z_tb <- as_tibble(zero_rev_col)
    # z_tb <- as.data.frame(zero_rev_col) # df is fine
  z_tb <- cbind(r = rownames(zero_rev_col), z_tb)
  
    ## Make a Column of Row Names, must be a factor.
      z_tb <- cbind("y" = make_ynames(zero), z_tb)
      
# 3. Transform the Data into a long format
z_long <- pivot_longer(z_tb, cols = names(z_tb)[-c(1,2)],
                       names_to = "x", 
                       values_to = "z")

  ## Super Important that column names are ordered Factors
        # Otherwise ggplot will rearrange
  z_long$x <- factor(z_long$x, levels = unique(z_long$x), ordered = TRUE)
    # x-value and column numbers will match 

  
# Print the Matrix
zero

# Make The Heatmap
ggplot(z_long, aes(x = x, y = y, fill = z)) +
  geom_tile()

  

```

###### Create a Function

Now that the matrix can be converted into a tidy data frame, we'll wrap that process into a function (with vim use `va{` , `y`, `p`, `va{:s/zero/mat/g<cr>`):



```{r}

matrix_as_long_tib <- function(mat) {
  
# 0. Name the Rows and Columns
  colnames(mat) <- make_colnames(mat)
  rownames(mat) <- make_rownames(mat)
    ## Row Names are reversed relative to y-axis values

# 1. Reverse the Columns
  mat_rev_col <- apply(mat, 2, rev)

# 2. Create a Data Frame
  z_tb <- as_tibble(mat_rev_col)
    # z_tb <- as.data.frame(mat_rev_col) # df is fine
  z_tb <- cbind(r = rownames(mat_rev_col), z_tb)
  
    ## Make a Column of Row Names, must be a factor.
      z_tb <- cbind("y" = make_ynames(mat), z_tb)
      
# 3. Transform the Data into a long format
z_long <- pivot_longer(z_tb, cols = names(z_tb)[-c(1,2)],
                       names_to = "x", 
                       values_to = "z")

  ## Super Important that column names are ordered Factors
        # Otherwise ggplot will rearrange
  z_long$x <- factor(z_long$x, levels = unique(z_long$x), ordered = TRUE)
    # x-value and column numbers will match 
  
  return(z_long)
}

(unsquare_test_matrix <- matrix(sample(c(TRUE, FALSE), replace = TRUE, size = 6^2), nrow = 9))
# (unsquare_test_matrix <- matrix(sample(c(TRUE, FALSE), replace = TRUE, size = 6^2), nrow = 6))
#(unsquare_test_matrix <- zero)


z_long <- matrix_as_long_tib(unsquare_test_matrix)
ggplot(z_long, aes(x = x, y = y, fill = z)) +
  geom_tile()
  

```

