---
title: "99 Final Exam"
tags: [Notebooks/Predictive Modelling]
date: 2018-June-4
author: "Ryan Greenup (1780 5315)"
# output:
#   rmarkdown::html_document:
#     theme: lumen
output:
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
    #Possible themes are cayman, tactile, architect, leonids, hpstr
---



#Preamble
##Clear Latent Variables
First clear latent variables that could be leftover from past work
```{r}
rm(list = ls())
```
##Load Packages
Now load all the necessary packages:
```{r}

if(require('pacman')){
    library('pacman')
  }else{
    install.packages('pacman')
    library('pacman')
  }
  
  pacman::p_load(ggmap, plotly, EnvStats, ggplot2, GGally, corrplot, dplyr, tidyr, stringr, reshape2, cowplot, ggpubr, reshape2, ggplot2, rmarkdown, dplyr, plotly, rstudioapi, wesanderson, RColorBrewer, colorspace, prettydoc, glmnet)
  
```

##Set Working Directory
It is not necessary to set the working directory because this is a markdown file, the working directory is automatically set to the location of file

##Load the data
the data can be loaded using a relative file path, because the working directory is automatically set to
the file location

```{r}
forest1.df <- read.csv(file = "forest1.csv", header = TRUE, ",")
forest2.df <- read.csv(file = "forest2.csv", header = TRUE, ",")
```

#(a+b) Compute the Column Means
Use a for loop to calculate the column means for the training data `forest1.df`
and automatically put these means over missing `NA` values in both the training
and testing data sets `forest1.df` and `forest2.df`.

It is important to use the training means (rather than) the testing means
because this can impede the performance of the model (due to data leakage)

```{r}


for (i in 1:ncol(forest1.df)) {
  if (is.numeric(forest1.df[,i])) {
    forest1.df[is.na(forest1.df[,i]),i] <- mean(forest1.df[,i], na.rm = TRUE)
  }
}

##Make sure to use the testing means not the training means
for (i in 1:ncol(forest2.df)) {
  if (is.numeric(forest2.df[,i])) {
    forest2.df[is.na(forest2.df[,i]),i] <- mean(forest1.df[,i], na.rm = TRUE)
  }
}


if(!(sum(is.na(forest1.df)) + sum(is.na(forest1.df)))){
  print("The cleaning was successfull")
}

```

Now export the cleaned data

```{r}
write.csv(x = forest1.df, file = "forest3.csv")
write.csv(x = forest2.df, file = "forest4.csv")
  
```

#(c) Compute the Standard Deviation and mean value for each column
In order to do this we need to remove the first column because it is
non-numeric, for the love of God remember that the numbers mightn't 
line up anymore.

```{r}

TrainMeanCol.vec <- sapply(X = forest1.df[,-1], FUN = mean)
TrainSDCol.vec <- sapply(X = forest1.df[,-1], FUN = sd)

data.frame("Column Names" = names(forest1.df[,-1]), "Mean Values" = TrainMeanCol.vec, "Standard Deviation" = TrainSDCol.vec) %>% print()

```

#(d) Standardising Data
The answer is the first one:

i. Divide each value by the standard deviation of the column, and subtract the column mean from the result


