---
title: "03 Multiple Linear Regression, Housing"
tags: [Notebooks/Predictive Modelling]
author: "Ryan Greenup"
date: "25 March 2018"
output: 
  html_document: 
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Preamble
First as a matter of housekeeping, we need to load in pertinent packages and 
set the working directory

```{r}
# Load Packages -----------------------------------------------------------

if(require('pacman')){
  library('pacman')
}else{
  install.packages('pacman')
  library('pacman')
}

pacman::p_load(xts, sp, gstat, ggplot2, rmarkdown, reshape2, ggmap,
               parallel, dplyr, plotly)


# Set the Working Directory -----------------------------------------------

set_wd <- function() {
  
  #Install the RStudio API package
  
  if(require('rstudioapi')){
    library('rstudioapi')
  }else{
    install.packages('rstudioapi')
    library('rstudioapi')
  }
  
  #Use the Rstudio API to get the working directory
  
  current_path <- getActiveDocumentContext()$path 
  setwd(dirname(current_path ))
  print( getwd() )
}

#set_wd()

# Import the Data Set -----------------------------------------------------
house.df <- read.csv("house2018.csv", header = TRUE, sep = ",")
house.df <- house.df[,c("price", "sqm", "bedrooms")]

    ##Create Assignments
    price <- house.df$price
    sqm   <- house.df$sqm
    broom <- house.df$bedrooms
  
```

#Inspect the Data

First we will Inspect the data:

```{r}
head(house.df)
```

##Use Plotting Tools

###Base Plot

A base plot can be used to get a rough idea for the distribution of the data: 
```{r}
layout(matrix(nrow = 1, 1:2))
plot(y = price, x = sqm, main = "House Size")
boxplot(price~broom, main = "Number of Rooms")
```

###GGplot

GGplot will allow us to look at all the data together:

```{r}
house.df$bedrooms <- as.factor(house.df$bedrooms)
ggplot(data = house.df, aes(x = sqm, y = price, col = bedrooms, shape = bedrooms)) +
  geom_point() +
  labs(col = "Bedrooms", shape = "Bedrooms") +
  xlab("Size (m^2)") +
  ylab("Price ($)") +
ggtitle("House Prices") +
  theme_classic()


```

##Discussion
By looking at the plots, it appears that both trends of price relative to these 
variables are linear,
thus geometrically it is likely the appropriate model is a plane.

Hence a muliple linear regression will be fitted to the data.


#Fit a model
A multiple linear regression can be fitted to the data thusly:
```{r}
#Remember that bedrooms MUST be a factor, NOT a continuous variable

houses.lm <- lm(price ~ sqm + broom)
summary(houses.lm)
```

Thus the best fitting multiple linear regression model is:

$$PRICE = 5362 \times SIZE + ROOMS \times 124239 - 65342$$
##Test the Model

```{r}
layout(matrix(nrow =2, 1:4))
plot(houses.lm)
houses.lm %>% resid() %>% shapiro.test()
```

The residuals of this model are slightly non-normal and the *Shapiro* p-value 
is greater than 0.1, thus this model may not be the appropriate one to use 

#Plot the Model
##Predict over a data frame
In order to plot this model, it is necessary to predict the model over a data frame.

```{r}
independent.df <- data.frame("sqm" = seq(from = 100, to = 200, length.out = 999),
                      "broom" = seq(from = 3, to = 3, length.out = 999))

#Remember bedrooms is a factor (i.e. discrete variable), that may or may not 
#need to be adjusted
#independent.df$broom <- as.factor(independent.df$broom)

price.mod.df <- predict(object = houses.lm, newdata = independent.df)
price.mod.df <- data.frame("Model.Price" = price.mod.df,
                           "Size" = independent.df)

head(price.mod.df)


```

##Create the Base Plots

```{r}

layout(matrix(nrow = 1, 1:2))
plot(y = price, x = sqm, 
     main = "House Size, 3 Bedrooms",
     xlab = "Size", ylab = "Price")
lines(y = price.mod.df$Model.Price,
      x = price.mod.df$Size.sqm)
plot(y = price.mod.df$Model.Price,
     x = price.mod.df$Size.sqm,
     pch = ".", main = "House Sizes, 3 Bedrooms",
     xlab = "size", ylab = "Price")


```

##Create GGPlots for better visualisation

```{r}
house.df.b3 <- subset(house.df, bedrooms == 3)

ggplot(data = house.df.b3, aes(x = sqm, y = price, shape = "b")) +
  geom_point(col = "limegreen", shape = 17, size = 2) +
  xlab("Size (m^2)") +
  ylab("Price ($)") +
ggtitle("House Prices, 3 Bedroom") +
  theme_classic() +
  geom_line(data = price.mod.df, mapping = aes(x = Size.sqm, y = Model.Price), col = "skyblue", lwd = 2)

```

#Conclusion
A multiple linear regression is a pretty poor model of house prices relative 
to size and number of bedrooms

This is demonstrated by observing the model at a cross section 
taken for 3 bedroom houses, the linear model only barely follows 
the data trend.
