---
title: "Final Exam"
date: 2018-June-4
tags: [Notebooks/Predictive Modelling]
author: "Ryan Greenup (1780 5315)"
# output:
#   rmarkdown::html_document:
#     theme: lumen
output:
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
    #Possible themes are cayman, tactile, architect, leonids, hpstr
---



#Preamble
##Clear Latent Variables
First clear latent variables that could be leftover from past work
```{r}
rm(list = ls())
```
##Load Packages
Now load all the necessary packages:
```{r}

if(require('pacman')){
    library('pacman')
  }else{
    install.packages('pacman')
    library('pacman')
  }
  
  pacman::p_load(ggmap, plotly, EnvStats, ggplot2, GGally, corrplot, dplyr, tidyr, stringr, reshape2, cowplot, ggpubr, reshape2, ggplot2, rmarkdown, dplyr, plotly, rstudioapi, wesanderson, RColorBrewer, colorspace, prettydoc, glmnet)
  
```

##Set Working Directory
It is not necessary to set the working directory because this is a markdown file, the working directory is automatically set to the location of file

##Load the data
the data can be loaded using a relative file path, because the working directory is automatically set to
the file location
```{r}

```

