---
title: "Titanic Survivors Plot"
tags: [Notebooks/Predictive Modelling]
output: html_notebook
---

#Preamble
##Load in the packages
```{r}

##Load Packages
if(require('pacman')){
  library('pacman')
}else{
  install.packages('pacman')
  library('pacman')
}

pacman::p_load(ggmap, plotly, EnvStats, ggplot2, GGally, corrplot, dplyr, tidyr,
               stringr, reshape2, cowplot, ggpubr, tidyverse, reshape2, ggplot2,
               rmarkdown, dplyr, plotly, rstudioapi, wesanderson, RColorBrewer,
               colorspace)

```

```{r}
#Load data
all.df <- read.csv("../0DataSets/titanictrain.csv", TRUE, ",")

```

#View the Data
```{r}
head(all.df)
```

#Create Plots
Consider how the passenger age and class affects the probability
of survival for a passenger, plots for this can be created thusly:

##Base Plot
In the Base package a scatter plot with continuous values or a 
boxplot with discrete values can be created thusly:
###Continous Scatter Plot
```{r}
plot(x = all.df$Pclass, y = all.df$Age, data = all.df, col = PlotCol.vec[all.df$Survived + 1], pch = 16, cex = 2)
```
###Categorical Box Plot
```{r}

###Create Categorical Factors
all.df$Pclass <- as.factor(all.df$Pclass)
all.df$Survived <- as.factor(all.df$Survived)
###Create the Plot (Box Plot)
PlotCol.vec <- rainbow_hcl(2)
plot(x = all.df$Pclass, y = all.df$Age, data = all.df,
     col = PlotCol.vec[all.df$Survived], pch = 16,
     xlab = "Passenger Class", ylab = "Passenger Age",
     main = "Titanic Survivors")
```
##ggplot2
GGplot may create more aesthetically pleasing plots:
```{r}

##Numeric
all.df <- read.csv("titanictrain.csv", TRUE, ",")
ggplot(all.df, aes(x = Pclass, y = Age, col = Survived)) +
  geom_point(lwd = 3) +
  guides(col = FALSE) +
  labs(title = "Titanic Survivors", x = "Passenger Class", y = "Passenger Age")
  #This is undesirable because the legend isn't done with discrete variables

##Factors

###Create Categorical Factors
all.df$Pclass <- as.factor(all.df$Pclass)
all.df$Survived <- as.factor(all.df$Survived)

ggplot(all.df, aes(x = Pclass, y = Age, col = Survived)) +
  geom_point(lwd = 4) +
  labs(title = "Titanic Survivors", x = "Passenger Class", y = "Passenger Age") +
  scale_color_discrete(labels = c("Perished", "Survived"), name = "Passenger Fate")
                      
```



