---
title: "Lecture 11 - R Code"
subtitle: "Creating Variogram Plots in R"
output: 
  html_notebook: 
    highlight: tango
    theme: spacelab
    toc: yes
  pdf_document: 
    toc: yes
  word_document: 
    toc: yes
editor_options: 
  chunk_output_type: inline
---
This is essentially a replication of the code used in the
lecture notes, The only difference is it is a little
less condenced and in HTML.

#Preamble
##Import Library

The following code will automatically install packages that are 
not already installed and load them into the library:

```{r}
if(require('pacman')){
  library('pacman')
}else{
  install.packages('pacman')
  library('pacman')
}

pacman::p_load(sp, EnvStats, gstat, ggplot2, rmarkdown)
```

##Create Assignments

Note that  Benthic.df is included in the *EnvStats* package and is not a 
data frame but an *sp* class, which is very similar to a data frame

```{r}
#Assignments
benthic <- Benthic.df
lat <- benthic$Latitude
lon <- benthic$Longitude
```


#Create a Variogram

```{r}
coordinates(benthic) = ~Longitude + Latitude
benthic.vg <- variogram(Index ~ 1, data = benthic)
```

##Create a base plot of that variogram

```{r}
plot(benthic.vg, main="Figure 11.1  Empirical variogram for Benthic Index")
```

##Create an attractive plot in ggplot

```{r}
ggplot(data = benthic.vg, aes(x = dist, y = gamma, col = gamma, size = np)) +
  geom_point() +
  labs(size = "Number of Point Pairs", col = "Sample Variogram Estimate",
       y = "Sample Variogram estimate", x = "Distance",
       title = 'Benthic Variogram Plot') +
  theme_bw() +
  geom_smooth(alpha = 0, size = 0.5)


```


#Fit a Variogram Model

There are three common models for variograms:

* Exponential
      + $C(h) = \sigma^2 \times e^{\frac{-h}{r}}$
* Gaussian
     + $C(h) = \sigma^2 \times e^{({\frac{-h}{r}})^2}$
* Spherical
     + $C(h) = \sigma^2 \times (1 - \frac{3h}{2r} + \frac{h^2}{2r^3})$
      
      
 
Unfourtunately `ggplot2` cannot take an object of class `variogramModel` and plot it 
(atleast I cant figure out how to do it), hence it is necessary to
use the base package to plot the model over the points.

It also does not appear possible to comine the plots using  `par(mfrow = c(1,3))`.

##Plot the model over the variogram
###Exponential
```{r}

benthic.vg.fit.exp <- fit.variogram(benthic.vg,
                                    model=vgm(1,"Exp", 0.5,1))
plot(benthic.vg, benthic.vg.fit.exp,
     main="Exponential variogram for Benthic Index")
```

###Gaussian
```{r}

benthic.vg.fit.gau <- fit.variogram(benthic.vg,
                                    model=vgm(1,"Gau", 0.5,1))
plot(benthic.vg, benthic.vg.fit.gau,
     main="Gaussian variogram for Benthic Index")
```

###Spherical
```{r}

benthic.vg.fit.sph <- fit.variogram(benthic.vg,
                                    model=vgm(1,"Sph", 0.5,1))
plot(benthic.vg, benthic.vg.fit.sph,
     main="Spherical variogram for Benthic Index")
```

#Directional Variograms

Because a directional variogram is still of the class `variogramModel`, 
it is not possible (to the best of my knowledge) to use `ggplot2`
in order to plot it.

The `alpha` option allows a vector of directional degrees to be specified, in this case, 0, 45, 90 and 135.

```{r}
benthic.dvg <- variogram(Index ~ 1,
                         data=benthic, alpha=c(0, 45,90,135))
benthic.dvg.fit <- vgm(1,"Sph", 0.5,1,anis=c(30,0.4))
plot(benthic.dvg, benthic.dvg.fit,
     main="Figure 11.5  Directional variograms for Benthic Index")
```

